---
title: "LAB Exercise-Jan8th"
author: "Luis Correia - 1006508566"
date: "9 January 2020"
output: 
  pdf_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


# Setup Tidyverse

Read in some packages that we'll be using:

```{r}
library(tidyverse)
```

Read in mortality rates for Canada, as seen in class. 

```{r}
path <- "C:/Users/LuisAlvaro/Documents/GitHub/applied-stats/data/CAN_Mx_1x1.txt"
dm <- read_table(path, skip = 2)
head(dm)
```

Repeat clean up as seen in class to make the ages and mortality rates numbers not characters.

```{r}
dm <- 
  dm %>% 
  mutate(Age = as.numeric(Age), 
         Female = as.numeric(Female), 
         Male = as.numeric(Male),
         Total = as.numeric(Male))
```

# Lab Exercises

1. Plot the ratio of male to female mortality rates over time and change the theme (e.g. `theme_bw()`)

First calculate the male/female ratio

```{r}
dm_ratio <- dm %>% 
  mutate(m_f_ratio = Male/Female)
head(dm_ratio)

```

... an then Plot graphics

```{r}
ggplot(dm_ratio, aes(Year, m_f_ratio, group = Year)) + 
  geom_line()+
  ylab("Mortality ratio")+
  theme_bw()+
  ggtitle("Mortality ratio of male/female over time")
```


2. Find the age that has the highest female mortality rate each year

```{r}
MaxAgeF <- 
  dm %>%
  group_by(Year) %>%
  mutate(max_FemMort = (Female == max(Female, na.rm = TRUE))) %>%
  filter(max_FemMort) 
MaxAgeF

#MaxAgeF2 <- 
#  dm %>%
#  group_by(Year, Age) %>%
#  mutate(max_FemMort = (Female == max(Female, na.rm = TRUE)))
#head(MaxAgeF2)

```


3. Use the `summarize_at()` function to calculate the standard deviation of mortality rates by age for the Male, Female and Total populations.

```{r}
SDev <- 
  dm %>%
  group_by(Age) %>%
  summarise_at(c("Female","Male","Total"),sd,na.rm = TRUE)
head(SDev)
```

