---
title: "STA2101-Applied Statistics I - Assignment #5"
author: "Luis Correia - Student No. 1006508566"
date: "21 de outubro de 2019"
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[CO,CE]{---STA2101-Applied Statistics I---}
- \fancyfoot[C]{Luis Correia - Student No. 1006508566}
- \fancyfoot[RO, LE] {\thepage}  
output: 
  pdf_document: default
  rmarkdown::pdf_document:
    fig_caption: yes
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, include=FALSE}
## All Packages used in this assignment

## Package for robust boxplots
require(robustbase)
require(xtable)
require(MASS)
require(MLmetrics)

```


# Question 7

People who raise large numbers of birds inhale potentially dangerous material, especially tiny fragments of feathers. Can this be a risk factor for lung cancer, controlling for other possible risk factors?  Which of those other possible risk factors are important? Here are the variables in the file \href{http://www.utstat.utoronto.ca/~brunner/data/illegal/birdlung.data.txt} 
{\texttt{http://www.utstat.utoronto.ca/$\sim$brunner/data/illegal/birdlung.data.txt}}.
These data are from a textbook called the \emph{Statistical Sleuth} by Ramsey and Schafer, and are used without permission.

```{r}
fpath <- "http://www.utstat.toronto.edu/~brunner/data/illegal/birdlung.data.txt"
DBirds <- read.table(fpath, header=FALSE)

```


\begin{center}
\begin{tabular}{|l|l|} \hline
\textbf{Variable} & \textbf{Values} \\ \hline
Lung Cancer                  & 1=Yes, 0=No \\ 
Gender                       & 1=Female, 0=Male \\ 
Socioeconomic Status         & 1=High, 0=Low \\ 
Birdkeeping                  & 1=Yes, 0=No \\ 
Age                          &  \\ 
Years smoked                 &  \\ 
Cigarettes per day           &  \\ \hline
\end{tabular}
\end{center}

If you look at \texttt{help(colnames)}, you can see how to add variable names to a data frame. It's a good idea, because if you can't remember which variables are which during the quiz, you're out of luck.

```{r}
colnames(DBirds) <- c("LungCanc", "Gender", "Status", "BirdKeep", "Age", "YearsSmoke", "CigPerDay")

DBirds$LungCanc <- factor(DBirds$LungCanc, levels = c(0, 1), labels = c("No", "Yes"))
DBirds$Gender <- factor(DBirds$Gender, levels = c(0, 1), labels = c("Male", "Female"))
DBirds$Status <- factor(DBirds$Status, levels = c(0, 1), labels = c("Low", "High"))
DBirds$BirdKeep <- factor(DBirds$BirdKeep, levels = c(0, 1), labels = c("No", "Yes"))
head(DBirds)
```


First, make tables of the binary variables using \texttt{table}, Use \texttt{prop.table} to find out the percentages. What proportion of the sample had cancer. Any comments?

```{r}
summary(DBirds)

# Make variables accessible
attach(DBirds)

# Finding the percentatge of cases with Cancer
TLCan <- addmargins(table(LungCanc), FUN = list(Total = sum)); TLCan
prop.table(table(LungCanc))

# Discovering the other percentages of data
table(LungCanc, Gender, Status, BirdKeep)

TBVar <- table(LungCanc, Gender, Status, BirdKeep)
prop.table(TBVar, margin = 1)
```

There is one primary issue in this study: Controlling for all other variables, is birdkeeping significantly related to the chance of getting lung cancer? Carry out a likelihood ratio test to answer the question.

(a) In symbols, what is the null hypothesis?

$H_0 : \beta_{BirdKeep=YES} = 0$

(b) What is the value of the likelihood ratio test statistic $G^2$? The answer is a number.
```{r}
fit1 <- glm(LungCanc~Gender+ Status + BirdKeep + Age + YearsSmoke + CigPerDay , family = binomial)
fit2 <- glm(LungCanc~Gender+ Status + Age + YearsSmoke + CigPerDay , family = binomial)

s.fit1 <- summary(fit1); s.fit1

s.fit2 <- summary(fit2); s.fit2

betahat1 <- fit1$coefficients; betahat1

betahat2 <- fit2$coefficients; betahat2

G2i <- fit2$deviance-fit1$deviance; G2i  # Difference between Full Model and Restricted(BirdKeep)

```

(c) What are the degrees of freedom for the test? The answer is a number.

```{r}

DF <- fit2$df.residual-fit1$df.residual; DF

```


(d) What is the $p$-value? The answer is a number.

```{r}
pSig <- 1-pchisq(G2i, df = DF)

cat("\n\np-Value = ", pSig)
```

(e) What do you conclude? Presence of a relationship is not enough. Say what happened.

```{r}
cat("\n\n The model suggests that people who keep birds have more chance to acquire LungCancer than other variables such as Age, Years of Smoke or Cigarettes per day")

```


(f) For a non-smoking, bird-keeping woman of average age and low socioeconomic status, what is the estimated probability of lung cancer? The answer (a single number) should be based on the full model.

```{r}

# AgeBar <- mean(Age[which(Gender=="Female")]); AgeBar  # Mean Age of Women

AgeBar <- mean(Age); AgeBar  # Mean Age of All persons (male & female)

x <- c(1,1,0,1,AgeBar,0,0); 

xb <- sum(x*fit1$coefficients)

phat <- exp(xb)/(1+exp(xb)); phat

cat("\n\nThe probability is ",phat,"\n\n")

head(DBirds)
xI <- data.frame(Gender="Female",Status="Low",BirdKeep="Yes",Age=AgeBar, YearsSmoke=0, CigPerDay=0); xI

cat("\n\n The Probability in easy way", predict(fit1,newdata = xI,type = "response"))

```


(g) Obtain a 95\% confidence interval for that last probability. Your answer is a pair of numbers. There is an easy way and a hard way. Do it the easy way.

```{r}
# Standard Error from the model
SEphat<-predict(fit1,newdata = xI,type="response",se.fit = TRUE); SEphat  #Std.Error = Std.Dev   

# Significance level (95% - two-sided)
pSig <- qnorm(0.975,mean=0,sd=1); pSig                      

# Lower and Upper bounts of CI for Theta
Lphat <- phat - pSig*SEphat$se.fit; Uphat <- phat + pSig*SEphat$se.fit     

# 95% signif level CI for pHat
cat("\nEstimated pHat = ",phat," and the 95 percent CI from [", Lphat,",",Uphat, "]\n\n")  

```


(h) Your answer to the last question made you uncomfortable. Why? Another approach is to start with a confidence interval for the log odds, and then use the fact that the function $p(x) = \frac{e^x}{1+e^x}$ is strictly increasing in $x$. Get the confidence interval this way. Again, your answer is a pair of numbers. Which confidence interval do you like more?

```{r}
Vhat1 <- vcov(fit1); Vhat1

P <- function (p) {
  return(exp(p)/(1+exp(p)))
}

# Female, Low, BirdKeep, Average Age, Non-Smoking
xT <- rbind(c(1,1,0,1,AgeBar,0,0))   

Sigma <- sqrt(xT%*%Vhat1%*%t(xT)); Sigma

#This is because xTbeta1 ~ N (xTbeta, xT.V.x)
xTbeta1 <- xT%*%betahat1

PI <- P(xTbeta1)

# Lower and Upper bounts of CI for PI
LPI <- P(xTbeta1 - pSig*Sigma); UPI <- P(xTbeta1 + pSig*Sigma)

# 95% signif level CI for PI
cat("\nEstimated PI = ",PI," and the 95 percent CI from [", LPI,",",UPI, "]\n\n")  

```


```{r}
message("\n\n VHat")

Vhat1 <- vcov(fit1); Vhat1

message("\n\n Inverse VHat")

VHatInv <- solve(Vhat1); VHatInv

```


(i) Naturally, you should be able to interpret all the $Z$-tests too. Which one is comparable to the main likelihood ratio test you have just done?

```{r}

```


(j) Controlling for all other variables, are the chances of cancer different for men and women?

```{r}

```


(k) Also, are \emph{any} of the explanatory variables related to getting lung cancer? Carry out a single likelihood ratio test. You could do it from the default output with a calculator, but use R. Get the $p$-value, too.

```{r}

G2k <- fit1$null.deviance-fit1$deviance; G2k  # Difference between Full Model and Restricted(BirdKeep)

DFk <- fit1$df.null-fit1$df.residual; DFk

pSig <- 1-pchisq(G2k, df = DFk)

cat("\n\np-Value = ", pSig)

```


(l) Now please do the same as the last item, but with a Wald test. Of course you should display the value of $W_n$, the degrees of freedom and the $p$-value.

```{r}
source("http://www.utstat.utoronto.ca/~brunner/Rfunctions/Wtest.txt")

# L0 <- rbind(c(0,1,1,1,1,1,1))  # H0: beta_i = 0 (i = 1, 2, 3, 4, 5, 6)

L0 <- rbind(c(0,1,0,0,0,0,0),
            c(0,0,1,0,0,0,0),
            c(0,0,0,1,0,0,0),
            c(0,0,0,0,1,0,0),
            c(0,0,0,0,0,1,0),
            c(0,0,0,0,0,0,1)
            )

Wtest(L0, betahat1, Vhat1)

```



(m) Finally and just for practice, fit a simple logistic regression model in which the single explanatory variable is number of cigarettes per day.

```{r}

fit3 <- glm(LungCanc~CigPerDay , family = binomial)

summary(fit3)

```


(\i) When a person from this population smokes ten more cigarettes per day, the odds of lung cancer are multiplied by $r$ (odds ratio). Give a point estimate of $r$. Your answer is a number.

```{r}

xCig <- c(1,10); # No. Cigarettes >= 10

xbCig <- sum(xCig*fit3$coefficients)

phatCig <- exp(xbCig)/(1+exp(xbCig)); phatCig

OddsCig <- phatCig/(1-phatCig); OddsCig

xNoCig <- c(1,0); # No. Cigarettes (No Smoker)

xbNoCig <- sum(xNoCig*fit3$coefficients)

phatNoCig <- exp(xbNoCig)/(1+exp(xbNoCig)); phatNoCig

OddsNoCig <- phatNoCig/(1-phatNoCig); OddsNoCig

OddsRatio <- OddsCig/OddsNoCig; OddsRatio

cat("\n\n The Odds of Lung Cancer for someone who smokes 10 cigarretes/day is ", OddsRatio)

```


(ii) Using the \texttt{vcov} function and the delta method, give an estimate of the asymptotic variance of $r$. Your answer is a number.

```{r}

vHat3 <- vcov(fit3); vHat3

xCigL <- c(0,10*exp(10*fit3$coefficients[2])); xCigL

gdot <- matrix(xCigL, nrow = 1, ncol=2); gdot


Sig3 <- gdot%*% vHat3 %*% t(gdot); Sig3

```


\textbf{Please bring your R printout for this question to the quiz.} Also, this question requires some paper and pencil work, and it would be fair to ask for something like that on the quiz too.



