---
title: "CHL5223 - Applied Bayesian Methods - Assignment 2"
author: "Luis Correia - Student No. 1006508566"
date: "February 22nd 2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
  rmarkdown::pdf_document:
    fig_caption: yes
    number_sections: yes
header-includes:
- \usepackage[margin=1in]{geometry} 
- \usepackage{amsmath,amsthm,amssymb,amsfonts}
- \usepackage{relsize}
- \usepackage{lscape}
- \usepackage{enumerate}
- \usepackage{enumitem}
- \usepackage{setspace}
- \usepackage{tikz}
- \usepackage{bm}
- \usepackage{bbm}
- \usepackage[utf8]{inputenc}
- \usepackage{mathtools, nccmath}
- \usepackage{fancyhdr}
- \usepackage{float}
- \usepackage{algorithm2e}
- \floatplacement{figure}{H}
- \floatplacement{table}{H}
- \pagestyle{fancy}
- \fancyhead[CO,CE]{---CHL5223 - Applied Bayesian Methods---}
- \fancyfoot[C]{Luis Correia - Student No. 1006508566}
- \fancyfoot[RO, LE] {\thepage}
- \setlength{\parskip}{1em}
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, fig.width=5, fig.height=4.0)
```

```{r, warning=FALSE, echo=FALSE}
library(tidyverse)
library(R2OpenBUGS)
library(kableExtra)
```

```{r}
library(ggplot2)
library(forecast)

# The palette with black - Used in Graphs with :
cbp1 <- c("#999999", "#E69F00", "#56B4E9", "#009E73",
          "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
cbp2 <- c("#000000", "#E69F00", "#56B4E9", "#009E73",
          "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
cbp3 <- c("#FFDB6D", "#C4961A", "#F4EDCA", "#D16103", 
          "#C3D7A4", "#52854C", "#4E84C4", "#293352")
```


\maketitle

\section{Question 1}

In the week 4 lecture, we looked at a regression example which predicted the log(brain weight) from the log(body weight) of different species of animals. In the lecture notes, an OpenBugs was used to create the sampling algorithm and then produce the sampling. In this homework, you are asked to create an MCMC simulations algorithm using basic R programming and the Gibbs sampler. The data are given in the file: \texttt{brnbdy.csv}. First, the probabilistic model and the conditional distributions are given, then the questions to be answered are given. Let $Y = log(\text{brain weight})$ and $X = log(\text{body weight})$ The probabilistic model for this problems is the following.

$$
Y_i|\alpha, \beta,X_i,\tau \sim \mathcal{N}(Y_i|\alpha+\beta X_i, \tau)
$$
$$
\alpha|\mu_{\alpha},\tau_{\alpha}\sim \mathcal{N}(\alpha|\mu_{\alpha},\tau_{\alpha})
$$
$$
\beta|\mu_{\beta},\tau_{\beta}\sim \mathcal{N}(\beta|\mu_{\beta},\tau_{\beta})
$$
$$
\tau|a_{\tau},b_{\tau}\sim Gamma(a_{\tau},b_{\tau})
$$

To finish off the model, assign the following values for the parameter variables: $\mu_{\alpha}=0=\mu_{\beta}$,$\tau_{\alpha}=0.0001=\tau_{\beta}$ and $a_{\tau}=0.0001=b_{\tau}$.

To construct the Gibbs sampler, you need to know the conditional distribution given all the other parameters in the model. (Note: I will only list the parameters that are needed in the below.) The conditional distribution for $\alpha$ given everything else is as follows:

$$
\alpha|X_{i=1:n},Y_{i=1:n},\beta,\tau,\mu_{\alpha},\tau_{\alpha}\sim\mathcal{N}(\mu_{\alpha}^{*},\tau_{\alpha}^{*})
$$
with

$$
\mu_{\alpha}^{*}=\dfrac{\tau\sum_{i=1}^n(Y_i-\beta X_i)+\tau_{\alpha}\mu_{\alpha}}{n\tau+\tau_{\alpha}}
$$
$$
\tau_{\alpha}^{*}=n\tau+\tau_{\alpha}
$$
The conditional distribution of $\beta$ given everything else is as follows:

$$
\beta|X_{i=1:n},Y_{i=1:n},\alpha,\tau,\mu_{\beta},\tau_{\beta}\sim\mathcal{N}(\mu_{\beta}^{*},\tau_{\beta}^{*})
$$
with

$$
\mu_{\beta}^{*}=\dfrac{\tau\sum_{i=1}^n[X_i(Y_i-\alpha)]+\tau_{\beta}\mu_{\beta}}{\tau\sum_{i=1}^nX_i^2+\tau_{\beta}}
$$
$$
\tau_{\beta}^{*}=\tau\sum_{i=1}^nX_i^2+\tau_{\beta}
$$

The conditional distribution of $\tau$ given everything else is as follows:

$$
\tau|X_{i=1:n},Y_{i=1:n},\alpha,\beta,a_{\tau},b_{\tau}\sim Gamma(a_{\tau}^{*},b_{\tau}^{*})
$$

with

$$
a_{\tau}^{*}=a_{\tau}+n/2\\
$$

$$
b_{\tau}^{*}=b_{\tau}+\frac{1}{2}\sum_{i=1}^n[Y_i-(\alpha+\beta X_i)]^2
$$
Do the following:

\subsection{item (a)}

Use R to run a Gibbs sampler algorithm, a Markov chain Monte Carlo algorithm, on this data. (DO NOT USE WinBugs, Openbugs, Jags or any of their relatives. The purpose is for you to do some basic programming.) You only need to run the MCMC for about $20,000$ iterations for this example. Note, that for full credit, you need to submit your code as part of the homework.

\medskip

{\setlength{\parindent}{0cm}\textit{Solution.}}

```{r}
# Read database
bb_data <- read.csv("Data/brnbdy.csv")

# Introducing Scale Power transformation - Weisberg(2013) pp.189-190
scalePower <- function (Z,Lambda=0, Mod = FALSE) {
  # Introduced by Box, Cox (1964) for strictly positive Z
  if (Mod)
    gm = exp(sum(log(Z)/length(Z))) 
  else
    gm = 1.0
  if (Lambda == 0)
    return(gm*log(Z))
  else
    return(gm^(1-Lambda)*(Z^Lambda-1)/Lambda)
}

# Setup Transformed Variables
Y  <-  scalePower(bb_data$brain)
X  <-  scalePower(bb_data$body)

```

For this question we will use the regression problem from week 4, where \texttt{log(Brain Weight)} (in grams) can be represented as a linear model from \texttt{log(Body Weight)} (in Kg).

From our data-set, it's easy to verify that the relationship we are about to model has linear characteristics as we can see from graph below:

```{r, fig.cap= 'Scatter-plot - log(Brain Weight) vs. log(Body Weight)', fig.height=3.0, fig.width=3.4}
# Prepare dataframe and plot brain vs. body weight
df <- data.frame (lBrW = Y, lBdW = X)

# Plot Graph for 
df %>% 
  ggplot(aes(x=X, y=Y))+
  geom_point(size=1.8)+
  labs(x = "log(Body Weight)", y = "log(Brain Weight)") +
  ggtitle(NULL)+
  theme_bw()

# Remove Temporary Variable 
remove(df)
```

\textbf{The Gibbs Sampling Algorithm}

The \textit{pseudo-code} of Gibbs Sampling algorithm applied to our problem can be summarized as follows:

\begin{algorithm}[H]
 set $N=20,000$\;
 set $a_{\tau}=b_{\tau}=0.0001$\; 
 set $\alpha^{(1)}=\beta^{(1)}=0$, $\tau^{(1)}=a_r/b_r$\; 
 \For{$i=2$ to N}{
    \textbf{Generate}  $\alpha^{(i)}$\;
    \Begin{
        Calculate $\tau_{\alpha}^{(i)}=n\tau^{(i-1)}+\tau_{\alpha}$\;
        Calculate $\mu_{\alpha}^{(i)}=\dfrac{\tau^{(i-1)}\sum_{j=1}^n(Y_j-\beta^{(i-1)} X_j)+\tau_{\alpha}\mu_{\alpha}}{\tau_{\alpha}^{(i)}}$\;
        Sample $\alpha^{(i)}$ from $\mathcal{N}(\mu_{\alpha}^{(i)},\tau_{\alpha}^{(i)})$\;
    }
    \textbf{Generate} $\beta^{(i)}$\;
    \Begin{
        Calculate $\tau_{\beta}^{(i)}=\tau^{(i-1)}\sum_{j=1}^nX_j^2+\tau_{\beta}$\;
        Calculate $\mu_{\beta}^{(i)}=\dfrac{\tau^{(i-1)}\sum_{j=1}^n[X_j(Y_j-\alpha^{(i)})]+\tau_{\beta}\mu_{\beta}}{\tau_{\beta}^{(i)}}$\;
        Sample $\beta^{(i)}$ from $\mathcal{N}(\mu_{\beta}^{(i)},\tau_{\beta}^{(i)})$\;
    }
    \textbf{Generate}  $\tau^{(i)}$\;
    \Begin{
        Calculate $a_{\tau}^{(i)}=a_{\tau}+n/2$\;
        Calculate $b_{\tau}^{(i)}=b_{\tau}+\frac{1}{2}\sum_{j=1}^n[Y_j-(\alpha^{(i)}+\beta^{(i)} X_j)]^2$\;
        Sample $\tau^{(i)}$ from $\Gamma(a_{\tau}^{(i)},b_{\tau}^{(i)})$\;
    }
 }
 \caption{Gibbs Sampling Algorithm}
\end{algorithm}

The corresponding \texttt{R-Code} is as follows: 

```{r, echo=TRUE}
# Setup Initial Variables
Mu_alpha <- Mu_beta <- 0
Tau_alpha <- Tau_beta <- 0.0001
a_Tau<- b_Tau <- 0.0001  # v.16

# Setup Iteration Variables
N <- 20000

alpha <- beta <- tau <- rep(N,0)

# Setup \theta^{(0)}
alpha[1] <- beta[1] <- 0
tau[1] <- 1

# Calculates the value of parameters, given everything else
sample_Alpha <- function (X, Y, beta, tau, mu_alpha, tau_alpha) {
  tau_alphaStar <- length(Y)*tau+tau_alpha
  mu_alphaStar <- (tau*sum(Y-beta*X)+tau_alpha*mu_alpha)/tau_alphaStar
  alpha <- rnorm(1,mean=mu_alphaStar, sd=sqrt(1/tau_alphaStar))
  return(alpha)
}

sample_Beta <- function (X, Y, alpha, tau, mu_beta, tau_beta) {
  tau_betaStar<- tau*sum(X^2)+tau_beta
  mu_betaStar <- (tau*sum(X*(Y-alpha))+tau_beta*mu_beta)/tau_betaStar
  beta <- rnorm(1,mean=mu_betaStar, sd=sqrt(1/tau_betaStar))
  return(beta)
}

sample_Tau <- function (X, Y, alpha, beta, a_r, b_r) {
  a_rStar <- a_r+length(Y)/2
  b_rStar <- b_r+0.5*sum((Y-(alpha+beta*X))^2)
  tau <- rgamma(1,shape=a_rStar, rate=b_rStar)
  return(tau)
}
    
Gibbs <- function(X, Y, alpha, beta, tau, N) {
  for (i in 2:N) {
    alpha[i] <- sample_Alpha(X, Y, beta[i-1], tau[i-1], Mu_alpha, Tau_alpha)
    beta[i] <- sample_Beta(X, Y, alpha[i], tau[i-1], Mu_beta, Tau_beta)
    tau[i] <- sample_Tau(X, Y, alpha[i], beta[i], a_Tau, b_Tau)   # v.17
  }
  return(list(Alpha = alpha, Beta = beta, Tau = tau))
}

# Set Seed for reproducibility
set.seed(737)

#Calculates 3 chains
ThetaC1 <- Gibbs(X, Y, alpha, beta, tau, N)
ThetaC2 <- Gibbs(X, Y, alpha, beta, tau, N)
ThetaC3 <- Gibbs(X, Y, alpha, beta, tau, N)

```

After running the algorithm we will have obtained $03$ chains of $20,000$ samples for posterior distributions of the parameters $\alpha$ (\texttt{Alpha[1:20000]}), $\beta$ (\texttt{Beta[1:20000]}) and $\tau$ (\texttt{Tau[1:20000]}) as requested.

For this question we also plotted the Auto-Correlation of estimates to verify its independence and set the \textit{burn-in} parameter to discard part of samples which may have some degree of interdependence.


```{r, fig.cap= "ACF plots - $\\alpha$, $\\beta$ and $\\tau$"}
# Plot ACF * ALPHA, BETA and TAU *

# Determine Burnin based on the convergence of ACF plot

# Build working dataframe
D_WorkAcfABT <- data.frame(ValCh1=ThetaC1$Alpha,
                           ValCh2=ThetaC1$Beta,
                           ValCh3=ThetaC1$Tau)


P1 <- ggAcf(D_WorkAcfABT$ValCh1, lag.max = 100)+
  labs(x = "Lag", y = expression(alpha)) +
  ggtitle(NULL)+
  theme_bw()
P2 <- ggAcf(D_WorkAcfABT$ValCh2, lag.max = 100)+
  labs(x = "Lag", y = expression(beta)) +
  ggtitle(NULL)+
  theme_bw()
P3 <- ggAcf(D_WorkAcfABT$ValCh3, lag.max = 100)+
  labs(x = "Lag", y = expression(tau)) +
  ggtitle(NULL)+
  theme_bw()

ggpubr::ggarrange(P1, P2, P3, ncol = 2, nrow = 2)

# Remove Working Variable to free memory
remove(D_WorkAcfABT)
```

The Auto-correlation plots (ACF) of the parameters indicate they are not correlated from the very initial lag's, so we can discard few estimates. For security, we will set the \textit{burn-in} parameter to $100$, which means we will discard the first $100$ estimates.  

```{r}
# Set Burn-in parameter
Burnin <- 100

# Prepare Data-Frame for next steps
D1_Work <- rbind(data.frame(alpha=ThetaC1$Alpha[Burnin:N],
                            beta=ThetaC1$Beta[Burnin:N],
                            tau=ThetaC1$Tau[Burnin:N],Chain=factor(1)),
                 data.frame(alpha=ThetaC2$Alpha[Burnin:N],
                            beta=ThetaC2$Beta[Burnin:N],
                            tau=ThetaC2$Tau[Burnin:N],Chain=factor(2)),
                 data.frame(alpha=ThetaC3$Alpha[Burnin:N],
                            beta=ThetaC3$Beta[Burnin:N],
                            tau=ThetaC3$Tau[Burnin:N],Chain=factor(3)))

```


```{r, fig.cap= "Traceplots of $\\alpha$, $\\beta$ and $\\tau$ and Scatterplot $\\alpha$ vs. $\\beta$"}
# Plot TracePlots * ALPHA, BETA and TAU *

set.seed(340)

Sz <- 5000

L <- N-Burnin+1

# Sampling 5000 points to generate "thinner" traceplots
S <- if (Sz <= L) sort(sample(Burnin:N, Sz, replace = FALSE)) else Burnin:N

# Build working dataframe
D_WorkTpAlpha <- data.frame(ValCh1=ThetaC1$Alpha[S],
                            ValCh2=ThetaC2$Alpha[S],
                            ValCh3=ThetaC3$Alpha[S])
P1 <- D_WorkTpAlpha %>%
  ggplot(aes(seq(from=1,to=Sz)))+
  geom_line(aes(y=ValCh1, colour=1), size=0.8)+ 
  geom_line(aes(y=ValCh2, colour=2), size=0.8)+ 
  geom_line(aes(y=ValCh3, colour=3), size=0.8)+ 
  labs(y = expression(alpha)) +
  theme_bw()

# Build working dataframe
D_WorkTpBeta <- data.frame(ValCh1=ThetaC1$Beta[S],
                           ValCh2=ThetaC2$Beta[S],
                           ValCh3=ThetaC3$Beta[S])
P2 <- D_WorkTpBeta %>%
  ggplot(aes(seq(from=1,to=Sz)))+
  geom_line(aes(y=ValCh1, colour=1), size=0.8)+ 
  geom_line(aes(y=ValCh2, colour=2), size=0.8)+ 
  geom_line(aes(y=ValCh3, colour=3), size=0.8)+ 
  labs(y = expression(beta)) +
  theme_bw()

# Build working dataframe
D_WorkTpTau <- data.frame(ValCh1=ThetaC1$Tau[S],
                          ValCh2=ThetaC2$Tau[S],
                          ValCh3=ThetaC3$Tau[S])
P3 <- D_WorkTpTau %>%
  ggplot(aes(seq(from=1,to=Sz)))+
  geom_line(aes(y=ValCh1, colour=1), size=0.8)+ 
  geom_line(aes(y=ValCh2, colour=2), size=0.8)+ 
  geom_line(aes(y=ValCh3, colour=3), size=0.8)+ 
  labs(y = expression(tau)) +
  theme_bw()

# Build working dataframe
D_WorkTpAlphaBeta <- data.frame(ValAlpha=c(ThetaC1$Alpha[S],
                                           ThetaC2$Alpha[S],
                                           ThetaC3$Alpha[S]),
                                ValBeta=c(ThetaC1$Beta[S],
                                          ThetaC2$Beta[S],
                                          ThetaC3$Beta[S]))

P4 <- D_WorkTpAlphaBeta %>%
  ggplot(aes(x=ValAlpha, y=ValBeta))+
  geom_point(size=0.6)+ 
  labs(x = expression(alpha), y = expression(beta)) +
  theme_bw()

ggpubr::ggarrange(P1+theme(axis.title.x=element_blank(),
                           legend.position="none"), 
                  P2+theme(axis.title.x=element_blank(),
                           legend.position="none"), 
                  P3+theme(axis.title.x=element_blank(),
                           legend.position="none"), 
                  P4+theme(legend.position="none"), ncol = 2, nrow = 2)

# Remove Working Variable to free memory
remove(D_WorkTpAlpha, D_WorkTpBeta, D_WorkTpTau, D_WorkTpAlphaBeta)
```

After discarding the burn-in elements, we can see from trace-plots\footnote{In order to avoid too many points on each graph, affecting the display performance of PDF file, we limited the number of points to $5,000$ which is a good representation of parameter's convergence.} that all chains are well mixed indicating the convergence of our estimates for our model's parameters.

We can also see from the joint distribution of $\alpha$ and $\beta$ that there is no evidence that parameters are correlated, we can move forward with our estimation for the posterior distributions.

\medskip

\subsection{item (b)}

Please give the posterior distribution (a plot and the summary statistics for the posterior distribution) for the parameters $\alpha$, $\beta$, and $\tau$. For the summary statistics of these posterior distributions it is sufficient to give the mean, standard deviation, and some type of $95\%$ credible region.

\medskip

{\setlength{\parindent}{0cm}\textit{Solution.}}

Using the samples generated by our algorithm for each parameter, we obtained the following graphs for posterior distributions of $\alpha$, $\beta$ and $\tau$.

```{r, fig.cap="Posterior Distributions $\\alpha$, $\\beta$ and $\\tau$"}
# Plot the Graphs of Alpha
P1 <- D1_Work %>%
  ggplot(mapping = aes(x = alpha, group = Chain))+
  geom_density(aes(colour=Chain), size=0.5)+
  labs(x = expression(alpha), y = "Density") +
  scale_color_manual(values = cbp2)+
  theme_bw()

# Plot the Graphs of Beta
P2 <- D1_Work %>%
  ggplot(mapping = aes(x = beta, group = Chain))+
  geom_density(aes(colour=Chain), size=0.5)+
  labs(x = expression(beta), y = "Density") +
  scale_color_manual(values = cbp2)+
  theme_bw()

# Plot the Graphs of Tau
P3 <-D1_Work %>%
  ggplot(mapping = aes(x = tau, group = Chain))+
  geom_density(aes(colour=Chain), size=0.5)+
  labs(x = expression(tau), y = "Density") +
  scale_color_manual(values = cbp2)+
  theme_bw()

ggpubr::ggarrange(P1, P2, P3,ncol = 1, nrow = 3)

```

... and the respective \textit{Summary Statistics} from our posterior distributions are as follows:


```{r}
# Summary Statistics - alpha
prt_SummaryStat <- function (sample_, parameter) {
  m_ <- mean(sample_)
  sd_ <- sd(sample_)
  cat("\n---- Summary Statistics for ",parameter," -----")
  cat("\nMean",parameter,": ",format(m_, digits = 2, nsmall = 3)) 
  cat("\nStd.Dev.",parameter,": ",format(sd_, digits = 2, nsmall = 4))
  CI <- qnorm(p = c(0.025,0.975), mean = m_, sd = sd_)
  # cat("\nC.I. for",parameter,": [",CI[1],",",CI[2],"]\n")
  cat("\n95% Credible Region for",parameter,"(", format(CI[1], digits = 2, nsmall = 3), 
    ",",format(CI[2], digits = 4, nsmall = 3),")\n")
}
prt_SummaryStat(D1_Work %>% pull(alpha), "alpha")
prt_SummaryStat(D1_Work %>% pull(beta), "beta")
prt_SummaryStat(D1_Work %>% pull(tau), "tau")

```

\medskip

\subsection{item (c)}

For an animal species with an average body weight of $55$, provide the distribution which represents your belief of the brain weight. (Note, one might consider two different values here. Basically, I want the estimate of brain weight as a function of (\{$\alpha$, $\beta$, body weight $=55$\}). Since this is a function of the random variables $\alpha$ and $\beta$, the estimate is also a random variable.) Provide the distribution by giving an estimate of the density (a plot) and summary statistics of the distribution (mean, standard deviation, and credible interval).

\medskip

{\setlength{\parindent}{0cm}\textit{Solution.}}

For an animal with body weight of $55$ Kg, we will use our samples of $\alpha$'s and $\beta$'s with $X=\texttt{log}(55)$ to generate samples of this animal. 

By doing so, the resulting distribution that represents our belief of brain weight for such animal is as follows:

```{r, fig.cap="Posterior Distribution of Brain Weight of an animal with 55 Kg"}
# Using the Alphas and Betas found on item (a) We will generate a sample of exp(Y_55)
D2_Work <- D1_Work %>% 
  mutate(Y_55 = alpha+beta*log(55))

D2_Work %>%
  ggplot(mapping = aes(x = exp(Y_55), group = Chain))+
  geom_density(aes(colour=Chain), size=0.8)+
  labs(x = expression("Estimated Brain Weight | Body Weight=55Kg"), y = "Density") +
  scale_color_manual(values = cbp2)+
  theme_bw()

```

```{r}
# Summary Statistics for the Distribution of an animal with average bodyweight of 55
m_a <- mean(exp(D2_Work$Y_55))
sd_a <- sd(exp(D2_Work$Y_55))
cat("\n---- Summary Statistics for Brain Weight | Body Weight = 55kg -----")
cat("\nMean : ",format(m_a, digits = 2, nsmall = 3)) 
cat("\nStd.Dev. : ",format(sd_a, digits = 2, nsmall = 4))
CI_a <- qnorm(p = c(0.025,0.975), mean = m_a, sd = sd_a)
cat("\n95% Credible Region (", format(CI_a[1], digits = 2, nsmall = 3), 
    ",",format(CI_a[2], digits = 2, nsmall = 3),")\n")
cat("\n>>> From this model, our belief of Brain Weight for an animal with \n")
cat("    average Body Weight of 55Kg is",format(m_a, digits = 2, nsmall = 2),
    "grams, with a 95% credible\n")
cat("    region of (",format(CI_a[1], digits = 2, nsmall = 2),",",
    format(CI_a[2], digits = 2, nsmall = 2),") grams.\n")

```


\pagebreak

\section{Question 2}

The following table gives the number of lung cancers for different age groups and different histories of smoking. Also, given is the number of person years where a potential cases was exposed (was eligible to be identified by the study as having lung cancer) and was in a particular age group-smoker class. (Data is from page $495$, Selvin, \textit{Practical Biostatistical Methods}, 1995, (Belmont, CA: Wadsworth).) Do the following with this data.

\subsection{item (a)}

Do a Bayesian Poisson regression on this data. Model the logarithm of the rate of deaths (per person-years) as a linear function of age and smoking category. Model both the smoking category and the age groupings as discrete variables. Please give the \texttt{WinBugs} code for this data. Provide the uni-variate posterior distributions of each level of smoking and age and also give the posterior distribution of the precision used in your model. (If you wish, you may report the "standard deviations" instead of the precision.)

\medskip

{\setlength{\parindent}{0cm}\textit{Solution.}}

```{r}
# Setup Data-set - Read Data
shist_data <- read.table(file="data/SmokeAgeDeath.csv",header=TRUE,sep=",")
colnames(shist_data) <- c("Smoking", "Age", "Deaths", "PersonYears")

# Setup Variables
NSmok=nrow(shist_data)  # Number of Items in data-set
NCSmok=4   # Number of classes of Smoking
NCAge=5    # Number of Classes of Age

# Setup OpenBugs running parameters
NSim <- 30000   # No. of simulations for productio
NChain <- 5     # No. of chains for production
NThin <- 5      # n.thin parameter for production
Burnin <- 10000 # Burn-In parameter for production
Sz <- 5000      # Size of samples for trace/acf plots

```

The data-set used for this question is listed below.

```{r}
# Printing the Data-Frame
shist_data %>%
  kbl(booktabs = TRUE,
      caption = "Deaths by Lung Cancer, per Age, Smoking Level and Person-Years") %>% 
  kable_styling(latex_options = "striped")
 
```

For this problem, lets define $X_i:= \text{Number of deaths by lung cancer}$ and $p_i:= \text{Person-Years}$, then our model can be represented as a Poisson Regression, as follows:

\begin{align*}
    \mathbb{E}(X_i)&=\big(\text{Death-rate}\big)\times\big(\text{Person-Years}\big)\\
    \implies\log\big(X_i\big)&=\log\big(\text{Death-Rate}\big)+\log\big(p_i\big)\\
    &=\Big(\beta_0+\sum_{j=1}^{Q_a}\sum_{k=1}^{Q_s}\beta_{jk}X_{i,j,k}\Big)+\log\big(p_i\big)
\end{align*}

with $Q_a=4$ is the number of categories of \textit{Age} and $Q_s=5$ is the number of categories of \textit{Smoking}

We can configure this model in \texttt{OpenBugs} with the following code:

\begin{verbatim}
model{
for(i in 1:NSmok)
{
#Age   Smoking   Deaths  PersonYears

Deaths[i]~dpois(lambda[i])
log(lambda[i]) <-  log(PersonYears[i]) + beta0 + beta.Age[Age[i]]+ beta.Smoking[Smoking[i]] + bias[i] 
bias[i] ~dnorm(0,tau)
bias.adj[i] <- bias[i] - mean(bias[]) 
}

# Loop for Age
for(ia in 1:NCAge){         
beta.Age[ia]~dnorm(0,tau.Age)
beta.Age.adj[ia] <- beta.Age[ia] - mean(beta.Age[]) 
}

# Loop for Smoking Level
for(is in 1:NCSmok){         
beta.Smoking[is]~dnorm(0,tau.Smoking)
beta.Smoking.adj[is] <- beta.Smoking[is] - mean(beta.Smoking[])
}

# Initial Priors - Verify
beta0 ~ dnorm(0, 0.008)   # Overall rate of disease
beta0.adj <- beta0 + mean(bias[]) + mean(beta.Age[])+ mean(beta.Smoking[])

# Initial Priors - Verify
std ~ dunif(0, 4)
tau <- 1/pow(std,2)

# Initial Priors - Verify
std.Age ~dunif(0, 4)
tau.Age <- 1/pow(std.Age,2)
std.Smoking ~ dunif(0,4) 
tau.Smoking <- 1/pow(std.Smoking,2)
}
\end{verbatim}


```{r}
cat("
model{
for(i in 1:NSmok)
{
#Age   Smoking   Deaths  PersonYears

Deaths[i]~dpois(lambda[i])
log(lambda[i]) <-  log(PersonYears[i]) + beta0 + beta.Age[Age[i]]+ beta.Smoking[Smoking[i]] + bias[i] 
bias[i] ~dnorm(0,tau)
bias.adj[i] <- bias[i] - mean(bias[]) 
}

# Loop for Age
for(ia in 1:NCAge){         
beta.Age[ia]~dnorm(0,tau.Age)
beta.Age.adj[ia] <- beta.Age[ia] - mean(beta.Age[]) 
}

# Loop for Smoking Level
for(is in 1:NCSmok){         
beta.Smoking[is]~dnorm(0,tau.Smoking)
beta.Smoking.adj[is] <- beta.Smoking[is] - mean(beta.Smoking[])
}

# Initial Priors - Verify
beta0 ~ dnorm(0, 0.008)   # Overall rate of disease
beta0.adj <- beta0 + mean(bias[]) + mean(beta.Age[])+ mean(beta.Smoking[])

# Initial Priors - Verify
std ~ dunif(0, 4)
tau <- 1/pow(std,2)

# Initial Priors - Verify
std.Age ~dunif(0, 4)
tau.Age <- 1/pow(std.Age,2)
std.Smoking ~ dunif(0,4) 
tau.Smoking <- 1/pow(std.Smoking,2)

}", file="LungCancerCentered.txt")

# Setup Parameters
params=c("beta.Age.adj",  "std.Age", "beta.Smoking.adj", "std.Smoking", "beta0.adj", "std", "tau", "tau.Age", "tau.Smoking")
```


```{r}
# Setup Initial Values
init.fun=function(){list(
  beta.Age=rnorm(5), std.Age=runif(1,0,4),
  beta.Smoking=rnorm(4), std.Smoking=runif(1,0,4), 
  std=runif(1,0,4), beta0=rnorm(1),
  bias=rnorm(20,0,.1))}

```


```{r, cache=TRUE}
# Run Open Bugs
set.seed(2602)
model.data <- list("Age", "Smoking","Deaths","PersonYears","NSmok", "NCSmok", "NCAge")
attach(shist_data)
LungCanc_v0=bugs(model.data, init.fun, params, model.file="LungCancerCentered.txt",
                 working.directory = ".", n.chains=NChain, n.iter=NSim, n.burnin=Burnin, n.thin=NThin)
detach(shist_data)

```

After simulating $N=30,000$ samples with \textit{burn-in} of $10,000$ samples, with $5$ chains the mean, standard-deviation and respective credible regions for model's parameters are listed in the following Report Summary. 

```{r}
# Get Simulation from OpenBugs
SArray= LungCanc_v0$sims.array   # Data Arrays
vname=attr(SArray,"dimnames")[3][[1]]  # Variable Names

```


```{r}
# Print Summary statistics of parameters oif Interest
RNames <- rownames(LungCanc_v0[["summary"]][,c(1:3,5,7)]) # List of parameters
df_Prt <- data.frame(Parameter = RNames)
df_Prt <- cbind(df_Prt, as_tibble(LungCanc_v0[["summary"]][,c(1:3,5,7)]))
df_Prt %>%
  kbl(booktabs = TRUE, 
      caption = "OpenBugs Summary - Deaths from Lung Cancer") %>% 
  kable_styling(latex_options = "striped")
```

\pagebreak

\textbf{Posterior Distribution of $\beta_0$}

```{r, fig.cap="Posterior Distributions of $\\beta_0$ (adjusted)"}
# Plot Sample densities to Compare Distributions over *BETA-0*
D_WorkBeta0 <- as_tibble(data.frame(LvlBeta=as.vector(SArray[,,"beta0.adj"])))

D_WorkBeta0 %>%
  ggplot(aes(x = LvlStd))+
  geom_density(aes(x=LvlBeta), size=1.2)+ 
  labs(x = expression(list(beta[0])), y = "Density") +
  scale_color_manual(values = cbp2)+
  theme_bw()

# Remove Working Variable to free memory
remove(D_WorkBeta0)

```

\textbf{Comments}:- The parameter $\beta_0$ on our model would represents the prediction of someone die from lung cancer with simultaneously both  categories of \textit{Age} and \textit{Smoke} equals null. Intuitively, this parameter could then represent the inherent propensity of any individual dying from lung cancer which certainly would contribute for the death-rate of our model. On the other hand, \textit{mathematically thinking} over our model, this individual would also be out of scope of our study since he/she doesn't fit any of the categories of Age and Smoking defined. For this reason we will not try to interpret the constant term as suggested by \cite{GelmanHill}.

\pagebreak

\textbf{Posterior Distribution of each level of Age}

```{r, fig.cap="Posterior Distributions of $\\beta_{Age}$ for levels 1-5 (adjusted)"}
# Plot Sample densities to Compare Distributions over *AGE*
D_WorkAge <- as_tibble(rbind(data.frame(LvlAge=as.vector(SArray[,,paste0("beta.Age.adj[", 1, "]")]),Age="0-45"),
                               data.frame(LvlAge=as.vector(SArray[,,paste0("beta.Age.adj[", 2, "]")]),Age="45-54"),
                               data.frame(LvlAge=as.vector(SArray[,,paste0("beta.Age.adj[", 3, "]")]),Age="55-64"),
                               data.frame(LvlAge=as.vector(SArray[,,paste0("beta.Age.adj[", 4, "]")]),Age="65-74"),
                               data.frame(LvlAge=as.vector(SArray[,,paste0("beta.Age.adj[", 5, "]")]),Age="74+")))


P0 <- D_WorkAge %>%
  ggplot(aes(x = LvlAge))+
  geom_density(aes(x=LvlAge,group=Age, colour=Age), size=1.2)+ 
  labs(x = expression(list(beta[Age_][i])), y = "Density") +
  scale_color_manual(values = cbp2)+
  theme_bw()

P0 + geom_segment(aes(x = -2, y = 2, xend = 1.5, yend = 4), size = 0.3,
                   arrow = arrow(length = unit(0.5, "cm")))

# Remove Working Variable to free memory
remove(D_WorkAge)

```

\textbf{Comments}:- We can see a \textit{increasing pattern} on the posterior distributions of $\beta_{Age}$ depending on age range. For younger people, the contribution of this parameter for death rate by lung cancer is less than 1 (i.e, for negative values for $\beta$'s) and increasing from age range 55+ years old on. This may suggest that people who are in in smaller age-range have less probability of dying by lung cancer than people who are on greater age-ranges.

\pagebreak

\textbf{Posterior Distribution of each level of Smoking}

```{r, fig.cap="Posterior Distributions of $\\beta_{Smoking}$ for levels 1-4 (adjusted)"}
# Plot Sample densities to Compare Distributions over *SMOKE LEVEL*
D_WorkSMoke <- as_tibble(rbind(data.frame(LvlSmoke=as.vector(SArray[,,paste0("beta.Smoking.adj[", 1, "]")]),Level="Lv1 - Never Smoked"),
                               data.frame(LvlSmoke=as.vector(SArray[,,paste0("beta.Smoking.adj[", 2, "]")]),Level="Lv2 - Past Smoker"),
                               data.frame(LvlSmoke=as.vector(SArray[,,paste0("beta.Smoking.adj[", 3, "]")]),Level="Lv3 - 1-20 Cig/Day"),
                               data.frame(LvlSmoke=as.vector(SArray[,,paste0("beta.Smoking.adj[", 4, "]")]),Level="Lv4 - +20 Cig/Day")))


D_WorkSMoke %>%
  ggplot(aes(x = LvlSmoke))+
  geom_density(aes(x=LvlSmoke,group=Level, colour=Level), size=1.2)+ 
  labs(x = expression(list(beta[Smoke_][i])), y = "Density") +
  scale_color_manual(values = cbp2)+
  theme_bw()

# Remove Working Variable to free memory
remove(D_WorkSMoke)

```

\textbf{Comments}:- From this comparative graph, we can see a clear pattern on the posterior distributions depending of level of smoking. For Levels 1 and 2 (Never or Past Smoker) the estimations of $\beta_{Smoke}$ are negative, implying that for those groups, the contribution of this parameter to death rate by lung cancer is less than 1. This may suggest that people who "never smoked" or are a "past smoker" have less probability of dying of lung cancer than people who are currently smoker. The reverse conclusion can also be drawn from people who are smokers, i.e. their smoking habit contributes to increase the probability of death by lung cancer.

\pagebreak

\textbf{Posterior Distribution of Standard Deviations / Precision of our model}

```{r, fig.cap="Posterior Distributions of Std. Deviation/Precision"}
# Plot Sample densities to Compare Distributions over *STD DEVIATION* and *PRECISION*
D_WorkStd <- as_tibble(rbind(data.frame(LvlStd=as.vector(SArray[,,"std.Age"]),Level="Age"),
                             data.frame(LvlStd=as.vector(SArray[,,"std.Smoking"]),Level="Smoking")))

D_WorkPrc <- as_tibble(rbind(data.frame(LvlStd=as.vector(SArray[,,"tau.Age"]),Level="Age"),
                             data.frame(LvlStd=as.vector(SArray[,,"tau.Smoking"]),Level="Smoking")))


P1 <- D_WorkStd %>%
  ggplot(aes(x = LvlStd))+
  geom_density(aes(x=LvlStd,group=Level, colour=Level), size=1.2)+ 
  labs(x = expression(list(sigma[Age]~","~sigma[Smoke])), y = "Density") +
  scale_color_manual(values = cbp2)+
  theme_bw()

P2 <- D_WorkPrc %>%
  ggplot(aes(x = LvlStd))+
  geom_density(aes(x=LvlStd,group=Level, colour=Level), size=1.2)+ 
  labs(x = expression(list(tau[Age]~","~tau[Smoke])), y = "Density") +
  scale_color_manual(values = cbp2)+
  theme_bw()

ggpubr::ggarrange(P1, P2, ncol = 1, nrow = 2)

# Remove Working Variable to free memory
remove(D_WorkStd, D_WorkPrc)

```

\medskip

\subsection{item (b)}

Provide a brief justification that your model has converged and that you have \textit{burned-in} your simulation enough. (Note: you may give \textit{thinned} plots of your simulations to avoid printing out plots which are suppose to represent several thousands or hundreds of thousand points. For the purpose of this question, you can provide simple diagnostics like trace/history plots and auto-correlation plots.)

\medskip

{\setlength{\parindent}{0cm}\textit{Solution.}}

For this question we will present \textit{thinned} trace and ACF plots of parameters estimates to demonstrate its \textit{convergence} and that they are \textit{uncorrelated}. 

For the sake of simplicity, ACF plots will contain only chain \#1 in order to avoid deteriorating the PDF's performance with uninformative data, since the results are quite identical for all chains.

Please also note that I used the \texttt{ggAcf()} function from package \texttt{forecast} due to its flexibility and configuration capabilities. The main difference of it from the standard \texttt{acf} is that \texttt{ggAcf()} starts plotting from $lag=1$, and \texttt{acf()} starts at $lag=0$. 

```{r, fig.cap= "Trace-plots of $\\beta_{Age}$ (adjusted)", fig.height=3.5, fig.width=4.5}
# Plot TracePlots * BETA.AGE *

# Sampling 5000 points to generate "thinner" traceplots
set.seed(312)

L <- NSim-Burnin

S <- if (Sz <= L) sort(sample(1:L, Sz, replace = FALSE)) else 1:Sz
 
# Build working dataframe

P <- list()  

for (i in 1:5) {
  # Build working dataframe
D_WorkTpAge <- data.frame(ValCh1=SArray[S,1,paste0("beta.Age.adj[", i, "]")],
                          ValCh2=SArray[S,2,paste0("beta.Age.adj[", i, "]")],
                          ValCh3=SArray[S,3,paste0("beta.Age.adj[", i, "]")],
                          ValCh4=SArray[S,4,paste0("beta.Age.adj[", i, "]")],
                          ValCh5=SArray[S,5,paste0("beta.Age.adj[", i, "]")])

p <- D_WorkTpAge %>%
        ggplot(aes(seq(from=1,to=Sz)))+
        geom_line(aes(y=ValCh1, colour=1), size=0.8)+ 
        geom_line(aes(y=ValCh2, colour=2), size=0.8)+ 
        geom_line(aes(y=ValCh3, colour=3), size=0.8)+ 
        geom_line(aes(y=ValCh4, colour=4), size=0.8)+ 
        geom_line(aes(y=ValCh5, colour=5), size=0.8)+ 
        labs(y = eval(bquote(expression(beta[Age_][.(i)])))) +
        theme_bw()

  P <- c(P, list(p))
}

ggpubr::ggarrange(P[[1]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"), 
                  P[[2]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[3]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[4]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[5]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"), ncol = 3, nrow = 2)

# Remove Working Variable to free memory
remove(p, P, D_WorkTpAge)
```


```{r, fig.cap= "Trace-plots of $\\beta_{Smoking}$ (adjusted)"}
# Plot TracePlots * BETA.SMOKING *

# Sampling 5000 points to generate "thinner" traceplots
set.seed(351)

L <- NSim-Burnin

S <- if (Sz <= L) sort(sample(1:L, Sz, replace = FALSE)) else 1:Sz
 
# Build working dataframe

P <- list()  

for (i in 1:4) {
  # Build working dataframe
  D_WorkTpSmok <- data.frame(ValCh1=SArray[S,1,paste0("beta.Smoking.adj[", i, "]")],
                             ValCh2=SArray[S,2,paste0("beta.Smoking.adj[", i, "]")],
                             ValCh3=SArray[S,3,paste0("beta.Smoking.adj[", i, "]")],
                             ValCh4=SArray[S,4,paste0("beta.Smoking.adj[", i, "]")],
                             ValCh5=SArray[S,5,paste0("beta.Smoking.adj[", i, "]")])
  
  p <- D_WorkTpSmok %>%
    ggplot(aes(seq(from=1,to=Sz)))+
    geom_line(aes(y=ValCh1, colour=1), size=0.8)+ 
    geom_line(aes(y=ValCh2, colour=2), size=0.8)+ 
    geom_line(aes(y=ValCh3, colour=3), size=0.8)+ 
    geom_line(aes(y=ValCh4, colour=4), size=0.8)+ 
    geom_line(aes(y=ValCh5, colour=5), size=0.8)+ 
    labs(y = eval(bquote(expression(beta[Smoke_][.(i)])))) +
    theme_bw()
  P <- c(P, list(p))
}

ggpubr::ggarrange(P[[1]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"), 
                  P[[2]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[3]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[4]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"), ncol = 2, nrow = 2)

# Remove Working Variable to free memory
remove(p, P, D_WorkTpSmok)
```

```{r, fig.cap= "Trace-plots of $\\beta_0$ (adjusted)"}
# Plot TracePlots * BETA.0 *

# Sampling 5000 points to generate "thinner" traceplots
set.seed(761)

L <- NSim-Burnin

S <- if (Sz <= L) sort(sample(1:L, Sz, replace = FALSE)) else 1:Sz
 
# Build working dataframe
D_WorkTpBeta0 <- data.frame(ValCh1=SArray[S,1,paste0("beta0.adj")],
                            ValCh2=SArray[S,2,paste0("beta0.adj")],
                            ValCh3=SArray[S,3,paste0("beta0.adj")],
                            ValCh4=SArray[S,4,paste0("beta0.adj")],
                            ValCh5=SArray[S,5,paste0("beta0.adj")])

P0 <- D_WorkTpBeta0 %>%
  ggplot(aes(seq(from=1,to=Sz)))+
  geom_line(aes(y=ValCh1, colour=1), size=0.8)+ 
  geom_line(aes(y=ValCh2, colour=2), size=0.8)+ 
  geom_line(aes(y=ValCh3, colour=3), size=0.8)+ 
  geom_line(aes(y=ValCh4, colour=4), size=0.8)+ 
  geom_line(aes(y=ValCh5, colour=5), size=0.8)+ 
  labs(x = "Iteration", y = expression(beta[0])) +
  theme_bw()

P0 + theme(axis.title.x=element_blank(),
           legend.position="none")

# Remove Working Variable to free memory
remove(D_WorkTpBeta0)
```


```{r, fig.cap= "Trace-plots of $\\sigma_{Age}$ and $\\sigma_{Smoking}$"}
# Plot TracePlots * STD.AGE and STD.SMOKING *

# Sampling 5000 points to generate "thinner" traceplots
set.seed(390)

L <- NSim-Burnin

S <- if (Sz <= L) sort(sample(1:L, Sz, replace = FALSE)) else 1:Sz
 
# Build working dataframe
D_WorkTpStdAge <- data.frame(ValCh1=SArray[S,1,paste0("std.Age")],
                             ValCh2=SArray[S,2,paste0("std.Age")],
                             ValCh3=SArray[S,3,paste0("std.Age")],
                             ValCh4=SArray[S,4,paste0("std.Age")],
                             ValCh5=SArray[S,5,paste0("std.Age")])

P1 <- D_WorkTpStdAge %>%
  ggplot(aes(seq(from=1,to=Sz)))+
  geom_line(aes(y=ValCh1, colour=1), size=0.8)+ 
  geom_line(aes(y=ValCh2, colour=2), size=0.8)+ 
  geom_line(aes(y=ValCh3, colour=3), size=0.8)+ 
  geom_line(aes(y=ValCh4, colour=4), size=0.8)+ 
  geom_line(aes(y=ValCh5, colour=5), size=0.8)+ 
  labs(x = "Iteration", y = expression(sigma[Age])) +
  theme_bw()

# Build working dataframe
D_WorkTpStdSmok <- data.frame(ValCh1=SArray[S,1,paste0("std.Smoking")],
                             ValCh2=SArray[S,2,paste0("std.Smoking")],
                             ValCh3=SArray[S,3,paste0("std.Smoking")],
                             ValCh4=SArray[S,4,paste0("std.Smoking")],
                             ValCh5=SArray[S,5,paste0("std.Smoking")])

P2 <- D_WorkTpStdSmok %>%
  ggplot(aes(seq(from=1,to=Sz)))+
  geom_line(aes(y=ValCh1, colour=1), size=0.8)+ 
  geom_line(aes(y=ValCh2, colour=2), size=0.8)+ 
  geom_line(aes(y=ValCh3, colour=3), size=0.8)+ 
  geom_line(aes(y=ValCh4, colour=4), size=0.8)+ 
  geom_line(aes(y=ValCh5, colour=5), size=0.8)+ 
  labs(x = "Iteration", y = expression(sigma[Smoke])) +
  theme_bw()

ggpubr::ggarrange(P1+theme(axis.title.x=element_blank(),
                           legend.position="none"), 
                  P2+theme(axis.title.x=element_blank(),
                           legend.position="none"), ncol = 1, nrow = 2)

# Remove Working Variable to free memory
remove(D_WorkTpStdAge, D_WorkTpStdSmok)
```

\textbf{Comment}:- From Trace Plots we can verify that all parameters estimations generated by our simulations \textit{converged}.

```{r, fig.cap= "ACF-plots of $\\beta_{Age}$ (adjusted, max-Lag=100)", fig.height=3.5, fig.width=4.5}
# Plot TracePlots * BETA.AGE *

P <- list()  

for (i in 1:5) {
  # Build working dataframe
  D_WorkAcfAge <- data.frame(ValCh1=SArray[,1,paste0("beta.Age.adj[", i, "]")])
  
  p <- ggAcf(D_WorkAcfAge$ValCh1, lag.max = 100)+
  labs(x = "Lag", y = eval(bquote(expression(beta[Age_][.(i)])))) +
    ggtitle(NULL)+
    theme_bw()
  P <- c(P, list(p))
}

ggpubr::ggarrange(P[[1]]+theme(axis.title.x=element_blank(),
                           axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 8),
                           legend.position="none"), 
                  P[[2]]+theme(axis.title.x=element_blank(),
                           axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 8),
                           legend.position="none"),
                  P[[3]]+theme(axis.title.x=element_blank(),
                           axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 8),
                           legend.position="none"),
                  P[[4]]+theme(axis.title.x=element_blank(),
                           axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 8),
                           legend.position="none"),
                  P[[5]]+theme(axis.title.x=element_blank(),
                           axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 8),
                           legend.position="none"), ncol = 3, nrow = 2)

# Remove Working Variable to free memory
remove(p, P, D_WorkAcfAge)
```

```{r, fig.cap= "ACF-plots of $\\beta_{Smoking}$ (adjusted, max-Lag=100)"}
# Plot ACF * BETA.SMOKING *

# Group 1-4 - Elaborate a List of Graphs

P <- list()  

for (i in 1:4) {
  # Build working dataframe
  D_WorkAcfSmok <- data.frame(ValCh1=SArray[,1,paste0("beta.Smoking.adj[", i, "]")])
  
  p <- ggAcf(D_WorkAcfSmok$ValCh1, lag.max = 100)+
  labs(x = "Lag", y = eval(bquote(expression(beta[Smoke_][.(i)])))) +
    ggtitle(NULL)+
    theme_bw()
  P <- c(P, list(p))
}

ggpubr::ggarrange(P[[1]]+theme(axis.title.x=element_blank(),
                               legend.position="none"), 
                  P[[2]]+theme(axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[3]]+theme(axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[4]]+theme(axis.title.x=element_blank(),
                               legend.position="none"), ncol = 2, nrow = 2)

# Remove Working Variable to free memory
remove(p, P, D_WorkAcfSmok)

```



```{r, fig.cap= "ACF-plots of $\\beta_0$ (adjusted, max-Lag=100)"}
# Plot TracePlots * BETA.0 *

# Build working dataframe
D_WorkAcfBeta0 <- data.frame(ValCh1=SArray[,1,paste0("beta0.adj")])

P1 <- ggAcf(D_WorkAcfBeta0$ValCh1, lag.max = 100)+
  labs(x = "Lag", y = expression(beta[0])) +
  ggtitle(NULL)+
  theme_bw()

P1+theme(axis.title.x=element_blank(),
         legend.position="none")

# Remove Working Variable to free memory
remove(D_WorkAcfBeta0)
```

\textbf{Comment}:- From ACF Plots we can also verify the parameters estimates generated by our simulations are \textit{uncorrelated}.

\subsection{item (c)}

Provide your belief as to the increase probability of death for someone who smokes > 20 cigarettes per day versus a nonsmoker. (That is, you would say that you believe that a smoker who smokes > 20 cigarettes per day has an increase risk of xxx times the risk of a nonsmoker.) Please, include both a "point" estimate of this increase risk and also state some interval estimate. Please identify the type of point and interval estimate that you are using.

\medskip

{\setlength{\parindent}{0cm}\textit{Solution.}}

In order to calculate the increase in probability for someone who smokes > 20 cigarettes per day versus a nonsmoker we will analyze the posterior mean of parameters $\beta_{Smoking}^{[1]}$ and $\beta_{Smoking}^{[4]}$, both presented in table below.

```{r}
# Comparison of Smoking Categories
SMokCat <- rbind(LungCanc_v0[["summary"]]["beta.Smoking.adj[1]",c(1:3,5,7)],
                 LungCanc_v0[["summary"]]["beta.Smoking.adj[4]",c(1:3,5,7)])
RNames <- c("beta.Smoking.adj[1]", "beta.Smoking.adj[4]")
df_Prt <- data.frame(Parameter = RNames)
df_Prt <- cbind(df_Prt, as_tibble(SMokCat))
df_Prt %>%
  kbl(booktabs = TRUE, 
      caption = "Smoking Categories - Never Smoked vs. >20 cigarettes per day") %>% 
  kable_styling(latex_options = "striped")

```

There are $02$ ways to answer this question.

Using the \textbf{modeling results} and \textbf{posterior estimates of the parameters}:

- The coefficient \textit{beta.Smoking.adj[1]} represents the expected difference in the number of deaths (on the logarithmic scale) if the individual "Never Smoked". Thus, the expected multiplicative increase is $e^{-0.4585647} = 0.6321904$.
- On the other side, individuals who smoke \textit{more than 20 cigarettes per day}, represented here by \textit{beta.Smoking.adj[4]}, have a multiplicative increase of $e^{0.9229055}=2.516592$.
- Comparing one group (never-smoker) against the other (>20 cigarettes per day), we can say that the second group has an increased risk of lung cancer of approximately $e^{0.9229055-(-0.4585647)}=3.98075$ times the risk of a individual who never smoked.

Another way of treat this question is by using \textbf{samples from $\beta_{Smoking[1]}$ and $\beta_{Smoking[4]}$} and then generate a \textit{sample of estimated increase in probability of death between both categories} by doing:

$$
\Delta p(\text{Death Rate})=\exp\big[\beta_{Smoking[4]}-\beta_{Smoking[1]}\big]
$$
From this sample is possible estimate the \textit{mean increase} and a corresponding \textit{credible interval} for our estimate.

By using this method, we have the following results: 

```{r}
# Calculate Probability Increase between categories
ProbInc <- exp(SArray[,,"beta.Smoking.adj[4]"]-SArray[,,"beta.Smoking.adj[1]"])
m_Smoke <- mean(ProbInc)
sd_Smoke <- sd(ProbInc)
cat("\n---- Summary Statistics for Increase in Probability -----")
cat("\nMean : ",format(m_Smoke, digits = 2, nsmall = 3)) 
cat("\nStd.Dev. : ",format(sd_Smoke, digits = 2, nsmall = 4))
CI_Smoking <- qnorm(c(0.025,0.975), mean=m_Smoke, sd=sd_Smoke)
cat("\n95% Credible Region (", format(CI_Smoking[1], digits = 2, nsmall = 3), 
    ",",format(CI_Smoking[2], digits = 2, nsmall = 3),")\n")
cat("\n>>> From this model, our belief of increase probability of death by lung cancer\n")
cat("    of someone who smokes >20 cigarettes per day vs. who never smoked is",format(m_Smoke, digits = 2, nsmall = 2))
cat("\n    with corresponding 95% credible region of (",    format(CI_Smoking[1], digits = 2, nsmall = 2),",",
    format(CI_Smoking[2], digits = 2, nsmall = 2),")\n")
```

We can see that both methods provide similar results, which confirms our approach.

\pagebreak

\section{Question 3}

(There is first preamble and background section. This is followed by the things to do for this problem.)

\textbf{Preamble and Background}:
First, there is a description of a data analysis problem which includes a data file and a model for the data. You are asked to fit the model to the data with a Bayesian analysis. Use Openbugs or Jags (or a similar program) to fit the model. Please show the model file used by the software program and provide some preliminary evidence that that MCMC has converged. (That is, it is sufficient to use such methods as trace plot examination and looking at a plot of the auto-correlation function.) Afterwards, there are additional questions about the model that you will need to answer.

The data in the CSV file named \texttt{DiabetesDrugEffect.csv} which contains data from 12 different clinical trial. The purpose of each of the clinical trials was to see if drug A was better than drug B. (Aside, they were looking to lower blood glucose in diabetic patients.). The outcome results were measured on a continuous scale. The column labeled "diff" was the main result in the study and is the overall average difference between drug A and drug B. The column "Sediff" is the estimate of the standard error of the difference estimate. The column "StudyN" is the number of subjects that were in each of the studies. The studies are labeled with the variable "StudyId" and take on values from 1 to 12. Note that the values from each study are obtained from the individual study publications. 

The goal is to combine the evidence in the 12 different studies. That is, the goal is to use the 12 different studies to obtain an overall belief in the average difference between the two drugs. In the medical literature, this is called a \textit{meta-analysis}. Here, we make the assumption that for each study site there are unknown site specific conditions so that each site has a different effect and that these individual effects can be modeled as a normal distribution around an overall average different. That is, the following model is assumed.

\begin{itemize}
    \item Define $\theta$ as the overall "true" effects of the difference between the two drugs. The main goal of this analysis is to learn about this parameter.
    \item Define $\delta_i$ be the true difference between treatments for the i-th study. These study effects are normally distributed around $\theta$ with a common variance of $\sigma_0^2$. The parameter $\sigma_0^2$ is of moderate interest.
    \item Let $Y_i$ be the observed difference for the i-th study. This is the estimate for $\delta_i$ from the original paper from the i-th study.
    \item Let $S_i$ be the observed standard error of the estimator $Y_i$ of $\delta_i$. It is assumed one can ignore the error in using the observed value $S_i$ for the true value. That is, here one is assuming that the observed $S_i$ is the exact value.
\end{itemize}

The implies of the following model which we can call model 1:

$$
Y_i|\delta_i,S_i\sim \text{Normal}(\text{mean}=\delta_i, \text{variance}=S_i^2)\\
$$
$$
S_i|\theta,\sigma_0^2\sim \text{Normal}(\text{mean}=\theta, \text{variance}=\sigma_0^2)
$$

Sometimes one is not concerned with the latent variables $\delta_i$. So, one can collapse the model into the following collapsed model which we call model 2:

$$
Y_i|\theta,\sigma_0^2\sim \text{Normal}(\text{mean}=\theta, \text{variance}=S_i^2+\sigma_0^2)
$$
(Note: in Openbugs/Jags, when one is using the normal distribution it is correct that the first parameter is still the mean but the second parameter is the precision which is 1/variance.)

\textbf{Do the following for question 3}

\subsection{item (a)}

Fit both model 1 and model 2 with Openbugs/Jags. Provide the model file. Provide evidence that the MCMC has converged. (You don't need to provide advance statistics. Trace plots and plot of the auto-correlation function with the appropriate discussion is sufficient for the purpose of this exam.) Also, provide summary statistics of the posterior distributions. (The statistics such as the posterior mean, posterior standard deviation, and credible region obtain from a print statement is sufficient.)

\medskip

{\setlength{\parindent}{0cm}\textit{Solution.}}

For this question, we generated $02$ \texttt{OpenBugs} configuration files, one for each model. 

They are listed below\footnote{The detailed R-Code can be verified at the Appendix of this document.}.

```{r}
# Setup - Read Data Diabetics
diab_data <- read.table(file="data/DiabetesDrugEffect.csv",header=TRUE,sep=",")
NDiab <- nrow(diab_data)

# Setup OpenBugs running parameters

NSim <- 30000   # No. of simulations for productio
NChain <- 5     # No. of chains for production
NThin <- 5      # n.thin parameter for production
Burnin <- 10000 # Burn-In parameter for production
Sz <- 5000      # Size of samples for trace/acf plots

```


```{r, cache=TRUE}
### **************** M O D E L  -  1 ****************** ###

### ---- SETUP MODELS ---- ###

# MODEL 1 - Splitted
cat("
model{
for(i in 1:NDiab)
{
#StudyID  StudyN  diff  Sediff

diff[i]~dnorm(delta[i], tau[i]) #v.19
delta[i]~dnorm(theta, tau0)
tau[i] <- 1/pow(Sediff[i],2)

}

# Initial Priors - Verify
theta ~dnorm(-1.5, 30)  # v.19
sigma0 ~dgamma(2, 1)    # v.19
tau0 <- 1/pow(sigma0,2)

}", file="DiabetDrug_M1.txt")

### ---- PARAM SETTINGS ---- ###

# Setup Parameters
paramsM1=c("theta", "tau0", "sigma0", "delta")

### ---- INITIALIZATION PROCEDURES ---- ###

# Setup Initial Values for Model 1
init.funM1=function(){list(
  theta = rnorm(1, mean=-1.5, sd=sqrt(.3)),
  sigma0=rgamma(1, 2, 1))}  # v.19

### ---- PROCESS MODELS ON OPENBUGS ---- ###

# Run Open Bugs - Model 1
set.seed(2602)
model1.data <- list("StudyID", "StudyN", "diff", "Sediff","NDiab")
attach(diab_data)
DiabetDrug_M1=bugs(model1.data, init.funM1, paramsM1, model.file="DiabetDrug_M1.txt",
                 working.directory = ".", n.chains=NChain, n.iter=NSim, n.burnin=Burnin, n.thin=NThin)
detach(diab_data)

### ---- ORGANIZE OUTPUTS ---- ###

# Get Simulation from OpenBugs - Model 1
DArrayM1= DiabetDrug_M1$sims.array   # Data Arrays
vname=attr(DArrayM1,"dimnames")[3][[1]]  # Variable Names

```


```{r, cache=TRUE}
### **************** M O D E L  -  2 ****************** ###

### ---- SETUP MODELS ---- ###

cat("
model{
for(i in 1:NDiab)
{
#StudyID  StudyN  diff  Sediff

diff[i]~dnorm(theta, tau[i])
tau[i] <- 1/(pow(sigma0,2)+pow(Sediff[i],2))
}

# Initial Priors - Verify
theta ~dnorm(-1.5, 30)  # v.19
sigma0 ~dgamma(2, 1)    # v.19

}", file="DiabetDrug_M2.txt")

### ---- PARAM SETTINGS ---- ###

# Setup Parameters
paramsM2=c("theta", "sigma0")


### ---- INITIALIZATION PROCEDURES ---- ###

# Setup Initial Values for Model 2
init.funM2=function(){list(
  theta = rnorm(1, mean=-1.5, sd=sqrt(.3)),
  sigma0=rgamma(1, 2, 1))}  # v.19

### ---- PROCESS MODELS ON OPENBUGS ---- ###

# Run Open Bugs - Model 2
set.seed(8421)
model2.data <- list("StudyID", "StudyN", "diff", "Sediff","NDiab")
attach(diab_data)
DiabetDrug_M2=bugs(model2.data, init.funM2, paramsM2, model.file="DiabetDrug_M2.txt",
                 working.directory = ".", n.chains=NChain, n.iter=NSim, n.burnin=Burnin, n.thin=NThin)
detach(diab_data)

### ---- ORGANIZE OUTPUTS ---- ###

# Get Simulation from OpenBugs - Model 2
DArrayM2= DiabetDrug_M2$sims.array   # Data Arrays
vname=attr(DArrayM2,"dimnames")[3][[1]]  # Variable Names

```


\subsubsection{Model 1 - Configuration file}

\begin{verbatim}
model{
for(i in 1:NDiab)
{
#StudyID  StudyN  diff  Sediff

diff[i]~dnorm(delta[i], tau[i])
delta[i]~dnorm(theta, tau0)
tau[i] <- 1/pow(Sediff[i],2)

}

# Initial Priors - Verify
theta ~dnorm(-1.5, 30)  # v.19
sigma0 ~dgamma(2, 1)    # v.19
tau0 <- 1/pow(sigma0,2)

}
\end{verbatim}

\subsubsection{Model 2 - Configuration file}

\begin{verbatim}
model{
for(i in 1:NDiab)
{
#StudyID  StudyN  diff  Sediff

diff[i]~dnorm(theta, tau[i])
tau[i] <- 1/(pow(sigma0,2)+pow(Sediff[i],2))
}

# Initial Priors - Verify
theta ~dnorm(-1.5, 30)  # v.19
sigma0 ~dgamma(2, 1)    # v.19

}
\end{verbatim}

\medskip

\subsubsection{Simulation and Summary Statistics for Models 1 and 2}

After simulating $N=30,000$ samples with \textit{burn-in} of $10,000$ samples, with $5$ chains the mean, standard-deviation and respective credible regions for model's parameters are listed in the following Report Summary.

```{r}
# Print Summary statistics of parameters oif Interest
RNames <- rownames(DiabetDrug_M1[["summary"]][,c(1:3,5,7)]) # List of parameters
df_Prt <- data.frame(Parameter = RNames)
df_Prt <- cbind(df_Prt, as_tibble(DiabetDrug_M1[["summary"]][,c(1:3,5,7)]))
df_Prt %>%
  kbl(booktabs = TRUE, 
      caption = "Model 1 - Difference between Drug A and B") %>% 
  kable_styling(latex_options = "striped")

```


```{r}
# Print Summary statistics of parameters oif Interest
RNames <- rownames(DiabetDrug_M2[["summary"]][,c(1:3,5,7)]) # List of parameters
df_Prt <- data.frame(Parameter = RNames)
df_Prt <- cbind(df_Prt, as_tibble(DiabetDrug_M2[["summary"]][,c(1:3,5,7)]))
df_Prt %>%
  kbl(booktabs = TRUE, 
      caption = "Model 2 - Difference between Drug A and B") %>% 
  kable_styling(latex_options = "striped")
```

\pagebreak

\subsubsection{Convergence of MCMC Simulation}

For this section, we will show the Trace and ACF plots to demonstrate the posterior estimates of the parameters converged and are uncorrelated, as well.

The same restriction applied on Question 2 was also applied, i.e., trace and ACF plots with number of points limited to $5,000$ to preserve performance of PDF file.

\underline{Convergence and Uncorrelation of Parameters from Model 1}

```{r, fig.cap= "Model 1 - Trace and ACF of $\\theta$"}
# Plot TracePlots and ACF of *THETA *

# Sampling 5000 points to generate "thinner" traceplots
set.seed(3961)

L <- NSim - Burnin

S <- if (Sz <= L) sort(sample(1:L, Sz, replace = FALSE)) else 1:Sz
 
# Build working dataframe
D_WorkTpTheta <- data.frame(ValCh1=DArrayM1[S,1,paste0("theta")],
                            ValCh2=DArrayM1[S,2,paste0("theta")],
                            ValCh3=DArrayM1[S,3,paste0("theta")],
                            ValCh4=DArrayM1[S,4,paste0("theta")],
                            ValCh5=DArrayM1[S,5,paste0("theta")])

P1 <- D_WorkTpTheta %>%
  ggplot(aes(seq(from=1,to=Sz)))+
  geom_line(aes(y=ValCh1, colour=1), size=0.8)+ 
  geom_line(aes(y=ValCh2, colour=2), size=0.8)+ 
  geom_line(aes(y=ValCh3, colour=3), size=0.8)+ 
  geom_line(aes(y=ValCh4, colour=4), size=0.8)+ 
  geom_line(aes(y=ValCh5, colour=5), size=0.8)+ 
  labs(y = expression(theta)) +
  theme_bw()

P2 <- ggAcf(DArrayM1[,1,paste0("theta")], lag.max = 100)+
  labs(x = "Lag", y = expression(theta)) +
  ggtitle(NULL)+
  theme_bw()

ggpubr::ggarrange(P1+theme(axis.title.x=element_blank(),
                           legend.position="none"), P2, ncol = 1, nrow = 2)

# Remove Working Variable to free memory
remove(D_WorkTpTheta)
```

```{r, fig.cap= "Model 1 - Trace and ACF of $\\sigma_0$"}
# Plot TracePlots * SIGMA0*

# Sampling 5000 points to generate "thinner" traceplots
set.seed(2391)

L <- NSim - Burnin

S <- if (Sz <= L) sort(sample(1:L, Sz, replace = FALSE)) else 1:Sz
 
# Build working dataframe
D_WorkTpSigma0 <- data.frame(ValCh1=DArrayM1[S,1,paste0("sigma0")],
                             ValCh2=DArrayM1[S,2,paste0("sigma0")],
                             ValCh3=DArrayM1[S,3,paste0("sigma0")],
                             ValCh4=DArrayM1[S,4,paste0("sigma0")],
                             ValCh5=DArrayM1[S,5,paste0("sigma0")])

P1 <- D_WorkTpSigma0 %>%
  ggplot(aes(seq(from=1,to=Sz)))+
  geom_line(aes(y=ValCh1, colour=1), size=0.8)+ 
  geom_line(aes(y=ValCh2, colour=2), size=0.8)+ 
  geom_line(aes(y=ValCh3, colour=3), size=0.8)+ 
  geom_line(aes(y=ValCh4, colour=4), size=0.8)+ 
  geom_line(aes(y=ValCh5, colour=5), size=0.8)+ 
  labs(y = expression(sigma[0])) +
  theme_bw()

P2 <- ggAcf(DArrayM1[,1,paste0("sigma0")], lag.max = 100)+
  labs(x = "Lag", y = expression(sigma[0])) +
  ggtitle(NULL)+
  theme_bw()

ggpubr::ggarrange(P1+theme(axis.title.x=element_blank(),
                           legend.position="none"), P2, ncol = 1, nrow = 2)

# Remove Working Variable to free memory
remove(D_WorkTpSigma0)
```

```{r, fig.cap= "Model 1 - Trace-plots for $\\hat{\\delta}_{1:4}$"}
# Plot TracePlots for Deltas

# Sampling 5000 points to generate "thinner" traceplots
set.seed(746)

L <- NSim - Burnin

S <- if (Sz <= L) sort(sample(1:L, Sz, replace = FALSE)) else 1:Sz

# Plot Function

#plotting <- function(d){
#  P <- list()
#  vars <- names(d)[!names(d)%in%c('channel','label')]
#  for (var in vars){
#    p <- ggplot(d, aes_string(x='channel', y=var)) + 
#      geom_boxplot(aes(fill=label)) + ggtitle(var)
#    P <- c(P, list(p))
#  }
#  return(list(plots=P, num=length(vars)))
#}

#PLOTS <- plotting(d)
#do.call(grid.arrange, c(PLOTS$plots, nrow = PLOTS$num))


# Group 1-4 - Elaborate a List of Graphs

P <- list()  

for (i in 1:4) {
  # Build working dataframe
  D_WorkTpDelta <- data.frame(ValCh1=DArrayM1[S,1,paste0("delta[", i, "]")],
                              ValCh2=DArrayM1[S,2,paste0("delta[", i, "]")],
                              ValCh3=DArrayM1[S,3,paste0("delta[", i, "]")],
                              ValCh4=DArrayM1[S,4,paste0("delta[", i, "]")],
                              ValCh5=DArrayM1[S,5,paste0("delta[", i, "]")])
  p <- D_WorkTpDelta %>%
    ggplot(aes(seq(from=1,to=Sz)))+
    geom_line(aes(y=ValCh1, colour=1), size=0.8)+ 
    geom_line(aes(y=ValCh2, colour=2), size=0.8)+ 
    geom_line(aes(y=ValCh3, colour=3), size=0.8)+ 
    geom_line(aes(y=ValCh4, colour=4), size=0.8)+ 
    geom_line(aes(y=ValCh5, colour=5), size=0.8)+ 
    labs(y = eval(bquote(expression(delta[.(i)])))) +
    theme_bw()
  P <- c(P, list(p))
}

ggpubr::ggarrange(P[[1]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"), 
                  P[[2]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[3]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[4]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"), ncol = 2, nrow = 2)

# Remove Working Variable to free memory
remove(p, P, D_WorkTpDelta)
```

```{r, fig.cap= "Model 1 - Trace-plots for $\\hat{\\delta}_{5:8}$"}
# Plot TracePlots for Deltas

# Sampling 5000 points to generate "thinner" traceplots
set.seed(456)

L <- NSim - Burnin

S <- if (Sz <= L) sort(sample(1:L, Sz, replace = FALSE)) else 1:Sz

# Group 5-8 - Elaborate a List of Graphs

P <- list()  

for (i in 5:8) {
  # Build working dataframe
  D_WorkTpDelta <- data.frame(ValCh1=DArrayM1[S,1,paste0("delta[", i, "]")],
                              ValCh2=DArrayM1[S,2,paste0("delta[", i, "]")],
                              ValCh3=DArrayM1[S,3,paste0("delta[", i, "]")],
                              ValCh4=DArrayM1[S,4,paste0("delta[", i, "]")],
                              ValCh5=DArrayM1[S,5,paste0("delta[", i, "]")])
  p <- D_WorkTpDelta %>%
    ggplot(aes(seq(from=1,to=Sz)))+
    geom_line(aes(y=ValCh1, colour=1), size=0.8)+ 
    geom_line(aes(y=ValCh2, colour=2), size=0.8)+ 
    geom_line(aes(y=ValCh3, colour=3), size=0.8)+ 
    geom_line(aes(y=ValCh4, colour=4), size=0.8)+ 
    geom_line(aes(y=ValCh5, colour=5), size=0.8)+ 
    labs(y = eval(bquote(expression(delta[.(i)])))) +
    theme_bw()
  P <- c(P, list(p))
}

ggpubr::ggarrange(P[[1]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"), 
                  P[[2]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[3]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[4]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"), ncol = 2, nrow = 2)

# Remove Working Variable to free memory
remove(p, P, D_WorkTpDelta)
```

```{r, fig.cap= "Model 1 - Trace-plots for $\\hat{\\delta}_{9:12}$"}
# Plot TracePlots for Deltas

# Sampling 5000 points to generate "thinner" traceplots
set.seed(198)

L <- NSim - Burnin

S <- if (Sz <= L) sort(sample(1:L, Sz, replace = FALSE)) else 1:Sz

# Group 9-12 - Elaborate a List of Graphs

P <- list()  

for (i in 9:12) {
  # Build working dataframe
  D_WorkTpDelta <- data.frame(ValCh1=DArrayM1[S,1,paste0("delta[", i, "]")],
                              ValCh2=DArrayM1[S,2,paste0("delta[", i, "]")],
                              ValCh3=DArrayM1[S,3,paste0("delta[", i, "]")],
                              ValCh4=DArrayM1[S,4,paste0("delta[", i, "]")],
                              ValCh5=DArrayM1[S,5,paste0("delta[", i, "]")])
  p <- D_WorkTpDelta %>%
    ggplot(aes(seq(from=1,to=Sz)))+
    geom_line(aes(y=ValCh1, colour=1), size=0.8)+ 
    geom_line(aes(y=ValCh2, colour=2), size=0.8)+ 
    geom_line(aes(y=ValCh3, colour=3), size=0.8)+ 
    geom_line(aes(y=ValCh4, colour=4), size=0.8)+ 
    geom_line(aes(y=ValCh5, colour=5), size=0.8)+ 
    labs(y = eval(bquote(expression(delta[.(i)])))) +
    theme_bw()
  P <- c(P, list(p))
}

ggpubr::ggarrange(P[[1]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"), 
                  P[[2]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[3]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[4]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"), ncol = 2, nrow = 2)

# Remove Working Variable to free memory
remove(p, P, D_WorkTpDelta)
```

```{r, fig.cap= "Model 1 - ACF-plots for $\\hat{\\delta}_{1:4}$"}
# Plot ACF plots for Deltas

# Sampling 5000 points to generate "thinner" traceplots

# Group 1-4 - Elaborate a List of Graphs

P <- list()  

for (i in 1:4) {
  # Build working dataframe
  D_WorkTpDelta <- data.frame(ValCh1=DArrayM1[,1,paste0("delta[", i, "]")])
  
  p <- ggAcf(D_WorkTpDelta$ValCh1, lag.max = 100)+
  labs(x = "Lag", y = eval(bquote(expression(delta[.(i)])))) +
    ggtitle(NULL)+
    theme_bw()
  P <- c(P, list(p))
}

ggpubr::ggarrange(P[[1]]+theme(axis.title.x=element_blank(),
                               legend.position="none"), 
                  P[[2]]+theme(axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[3]]+theme(axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[4]]+theme(axis.title.x=element_blank(),
                               legend.position="none"), ncol = 2, nrow = 2)

# Remove Working Variable to free memory
remove(p, P, D_WorkTpDelta)
```

```{r, fig.cap= "Model 1 - ACF-plots for $\\hat{\\delta}_{5:8}$"}
# Plot ACF plots for Deltas

# Sampling 5000 points to generate "thinner" traceplots

# Group 5-8 - Elaborate a List of Graphs

P <- list()  

for (i in 5:8) {
  # Build working dataframe
  D_WorkTpDelta <- data.frame(ValCh1=DArrayM1[,1,paste0("delta[", i, "]")])
  
  p <- ggAcf(D_WorkTpDelta$ValCh1, lag.max = 100)+
  labs(x = "Lag", y = eval(bquote(expression(delta[.(i)])))) +
    ggtitle(NULL)+
    theme_bw()
  P <- c(P, list(p))
}

ggpubr::ggarrange(P[[1]]+theme(axis.title.x=element_blank(),
                               legend.position="none"), 
                  P[[2]]+theme(axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[3]]+theme(axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[4]]+theme(axis.title.x=element_blank(),
                               legend.position="none"), ncol = 2, nrow = 2)

# Remove Working Variable to free memory
remove(p, P, D_WorkTpDelta)
```


```{r, fig.cap= "Model 1 - ACF-plots for $\\hat{\\delta}_{9:12}$"}
# Plot ACF plots for Deltas

# Group 9-12 - Elaborate a List of Graphs

P <- list()  

for (i in 9:12) {
  # Build working dataframe
  D_WorkTpDelta <- data.frame(ValCh1=DArrayM1[,1,paste0("delta[", i, "]")])
  
  p <- ggAcf(D_WorkTpDelta$ValCh1, lag.max = 100)+
  labs(x = "Lag", y = eval(bquote(expression(delta[.(i)])))) +
    ggtitle(NULL)+
    theme_bw()
  P <- c(P, list(p))
}

ggpubr::ggarrange(P[[1]]+theme(axis.title.x=element_blank(),
                               legend.position="none"), 
                  P[[2]]+theme(axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[3]]+theme(axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[4]]+theme(axis.title.x=element_blank(),
                               legend.position="none"), ncol = 2, nrow = 2)
# Remove Working Variable to free memory
remove(p, P, D_WorkTpDelta)
```

\underline{Convergence and Uncorrelation of Parameters from Model 2}

```{r, fig.cap= "Model 2 - Trace and ACF of $\\theta$"}
# Plot TracePlots and ACF of *THETA *

# Sampling 5000 points to generate "thinner" traceplots
set.seed(3871)

L <- NSim - Burnin

S <- if (Sz <= L) sort(sample(1:L, Sz, replace = FALSE)) else 1:Sz
 
# Build working dataframe
D_WorkTpTheta <- data.frame(ValCh1=DArrayM2[S,1,paste0("theta")],
                            ValCh2=DArrayM2[S,2,paste0("theta")],
                            ValCh3=DArrayM2[S,3,paste0("theta")],
                            ValCh4=DArrayM2[S,4,paste0("theta")],
                            ValCh5=DArrayM2[S,5,paste0("theta")])

P1 <- D_WorkTpTheta %>%
  ggplot(aes(seq(from=1,to=Sz)))+
  geom_line(aes(y=ValCh1, colour=1), size=0.8)+ 
  geom_line(aes(y=ValCh2, colour=2), size=0.8)+ 
  geom_line(aes(y=ValCh3, colour=3), size=0.8)+ 
  geom_line(aes(y=ValCh4, colour=4), size=0.8)+ 
  geom_line(aes(y=ValCh5, colour=5), size=0.8)+ 
  labs(y = expression(theta)) +
  theme_bw()

P2 <- ggAcf(DArrayM2[,1,paste0("theta")], lag.max = 100)+
  labs(x = "Lag", y = expression(theta)) +
  ggtitle(NULL)+
  theme_bw()

ggpubr::ggarrange(P1+theme(axis.title.x=element_blank(),
                           legend.position="none"), P2, ncol = 1, nrow = 2)

# Remove Working Variable to free memory
remove(D_WorkTpTheta)
```

```{r, fig.cap= "Model 2 - Trace and ACF of $\\sigma_0$"}
# Plot TracePlots * SIGMA0*

# Sampling 5000 points to generate "thinner" traceplots
set.seed(8463)

L <- NSim - Burnin

S <- if (Sz <= L) sort(sample(1:L, Sz, replace = FALSE)) else 1:Sz
 
# Build working dataframe
D_WorkTpSigma0 <- data.frame(ValCh1=DArrayM2[S,1,paste0("sigma0")],
                             ValCh2=DArrayM2[S,2,paste0("sigma0")],
                             ValCh3=DArrayM2[S,3,paste0("sigma0")],
                             ValCh4=DArrayM2[S,4,paste0("sigma0")],
                             ValCh5=DArrayM2[S,5,paste0("sigma0")])

P1 <- D_WorkTpSigma0 %>%
  ggplot(aes(seq(from=1,to=Sz)))+
  geom_line(aes(y=ValCh1, colour=1), size=0.8)+ 
  geom_line(aes(y=ValCh2, colour=2), size=0.8)+ 
  geom_line(aes(y=ValCh3, colour=3), size=0.8)+ 
  geom_line(aes(y=ValCh4, colour=4), size=0.8)+ 
  geom_line(aes(y=ValCh5, colour=5), size=0.8)+ 
  labs(y = expression(sigma[0])) +
  theme_bw()

P2 <- ggAcf(DArrayM2[,1,paste0("sigma0")], lag.max = 100)+
  labs(x = "Lag", y = expression(sigma[0])) +
  ggtitle(NULL)+
  theme_bw()

ggpubr::ggarrange(P1+theme(axis.title.x=element_blank(),
                           legend.position="none"), P2, ncol = 1, nrow = 2)

# Remove Working Variable to free memory
remove(D_WorkTpSigma0)
```

\textbf{Comment}:- From Trace and ACF Plots we can verify the parameters estimates generated by our simulations for \textbf{Models 1 and 2} \textit{converged} and are \textit{uncorrelated}.

\medskip

\subsection{item (b)}

For model 1 and model 2, compare the estimates of the posterior distribution of $\theta$ and $\sigma_0$ from the two models. Are they close?

\medskip

{\setlength{\parindent}{0cm}\textit{Solution.}}

\textbf{Posterior distributions of $\theta$ and $\sigma_0$ compared}

```{r, fig.cap= "Posterior Distributions of $\\theta$ and $\\sigma_0^2$ for Models 1 and 2"}
# Plot Sample densities to Compare Distributions of *THETA* and *SIGMA0*

set.seed(9023)

L <- NSim-Burnin

S <- if (Sz <= L) sort(sample(1:L, Sz, replace = FALSE)) else 1:Sz
 
D_WorkThetas <- as_tibble(rbind(data.frame(LvlStd=as.vector(DArrayM1[S,,paste0("theta")]), Model = "M1"),
                                data.frame(LvlStd=as.vector(DArrayM2[S,,paste0("theta")]), Model = "M2")))
D_WorkSigmas <- as_tibble(rbind(data.frame(LvlStd=as.vector(DArrayM1[S,,paste0("sigma0")]), Model = "M1"),
                                data.frame(LvlStd=as.vector(DArrayM2[S,,paste0("sigma0")]), Model = "M2")))


P1 <- D_WorkThetas %>%
  ggplot(aes(x = LvlStd))+
  geom_density(aes(x=LvlStd, group=Model, colour=Model), size=0.8)+ 
  labs(x = expression(list(theta[Model_][1]~"vs."~theta[Model_][2])), y = "Density") +
  scale_color_manual(values = cbp2)+
  theme_bw()

P2 <- D_WorkSigmas %>%
  ggplot(aes(x = LvlStd))+
  geom_density(aes(x=LvlStd, group=Model, colour=Model), size=0.8)+ 
  labs(x = expression(list((sigma[0])[Model_][1]~"vs."~(sigma[0])[Model_][2])), y = "Density") +
  scale_color_manual(values = cbp2)+
  theme_bw()

ggpubr::ggarrange(P1, P2, ncol = 1, nrow = 2)

# Remove Working Variable to free memory
remove(D_WorkThetas, D_WorkSigmas)

```

\begin{itemize}
    \item \textbf{Comparison of $\theta$} - Both densities share almost the same location and scale parameters, which makes their respective densities to be close. As we started from common priors for both models, the hyper-parameter $\theta$ will present similar estimates, as well.
    \item \textbf{Comparison of $\sigma_0$} - The estimates for standard deviations of $\delta$ is almost the same for both models. This is explained because we started from same priors for $\sigma_0$ for both models. Despite of the fact they are used in different ways to model the precision of the difference between the drugs, they perform quite well in both models. 
\end{itemize}

By this modelling exercise, we could verify that the posterior overall difference between drug A and drug B (modeled through $\theta$) are almost the same, and the condensed Model-2 is as good as complete Model-1. 

\subsection{item (c)}

Using model 1, one gets an estimate of the posterior distribution of the individual study effects, $\delta_i$. That is, one can get the posterior mean, standard deviation and credible region. Also, from the original, individual study data, one can use the variables "diff" and "Sediff" to get an estimate of the study effects from the original studied publications. Therefore, one can use these values to obtained the estimated effect from the study and the confidence interval for the estimate. Compare the estimates from the individual studies to the estimates from the meta-analysis. For each study compare the values of "diff" and the associated confidence interval with the posterior mean of the $\delta_i$ estimate and the associated credible region.
\begin{enumerate}[label=(\roman*)]
    \item You will probably find that the estimate of the individual study effect using "diff" will be different than the estimate using the posterior mean of $\delta$. When considering the movement of the estimate from the value "diff" to the posterior mean of $\delta$, what general direction is this movement?
    \item When considering the length of the confidence intervals (using the individual study publications) compared to the credible region (obtained from the Bayesian meta-analysis), what generally happens to the change in the length?
    \item When looking at study 2, note that this study had the largest value of "Sediff" which is the standard error of the estimate in the original study. You might notice that there is a large difference between "diff" compared to the posterior mean of $\delta$ compared to most of the other studies. What might be causing this large shift. (Look at some other features of this study.)
\end{enumerate}

\medskip

{\setlength{\parindent}{0cm}\textit{Solution.}}

\textbf{Comparison between "diff" / posterior means of $\delta_i$ and respective C.I. and C.R.'s}

In order to answer this question, we plotted $02$ comparative graphs composed by \textit{Credible Regions} and \textit{Confidence Intervals}, with its respective posterior means and estimates.  

```{r, fig.cap= "Comparative C.R. vs. C.I. for Model-1 and Studies", fig.width=4, fig.height=3.2}
# Prepare Data - Get delta[1-12]
D3_Work <- as.data.frame(DiabetDrug_M1[["summary"]][4:15,c(1:3,5,7)])

# Generates Studies labels
StudyLst <- rapply(list(diab_data$StudyID), sprintf, fmt = "%02d", how = "replace")

# Collects C.I's of observed values from studies
CI_Study <- data.frame(ID = StudyLst[[1]],   
                       lower = diab_data$diff-diab_data$Sediff, 
                       estim = diab_data$diff, 
                       upper = diab_data$diff+diab_data$Sediff, 
                       Source = "Study")

# Collects C.R's from Summary report 
CI_Estim <- data.frame(ID = StudyLst[[1]], 
                       lower = D3_Work$`2.5%`, 
                       estim = D3_Work$mean, 
                       upper = D3_Work$`97.5%`, 
                       Source = "Model-1")

# Organize data to generate graphs & analysis
D4_Work <- rbind(CI_Study, CI_Estim)

#  (i) Identifying the direction of change of estimates
# (ii) Identifying differences o length of CR and CI for delta and "diff"
P1 <- D4_Work %>% 
  ggplot() +
  geom_point(aes(x=ID, y=estim, color = Source, shape = Source, size = 0.5)) +
  geom_errorbar(aes(x=ID, ymin=lower, ymax=upper,
                    color = Source), width = 1) +
  xlab("Study No.") +
  ylab(expression("C.R.+"~delta[i]~" and 'diff'")) +
  scale_color_manual(values = cbp2)+
  facet_wrap(~Source)+
  theme_bw()

P1 + theme(legend.position="none",
           axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 8))
```


```{r, fig.cap= "Comparative Posterior Means vs. $\\delta$\'s for Model-1 and Studies", fig.width=4, fig.height=3.2}
# (i) Identifying the direction of posterior mean of delta and "diff"
P2 <- D4_Work %>% 
  ggplot() +
  geom_point(aes(x=ID, y=estim, colour = Source, shape = Source), size=5) +
  xlab(expression("Study No.")) +
  ylab(expression(delta[i]~" and 'diff'")) +
  scale_color_manual(values = cbp2)+
  theme_bw()

P2 + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 8))

remove(D3_Work, D4_Work, StudyLst)

```

(i) From these graphs we can see that for $1/3$ (i.e., for studies $1, 10, 11$ and $12$), the posterior mean for $\delta_i$ is smaller than the obtained in that study. For the other group (i.e, $2/3$ or the remaining $8$ studies), we have the reverse scenario - posterior mean is smaller than obtained in studies. If we consider that \textit{smaller negative numbers} of $\delta_i$ and "diff" represent larger differences between the two drugs, we can say our overall scenario is more \textit{conservative}, considering a more balanced view of the effect of both drugs. 

(ii) When analyzing the C.R's and C.I.'s, we observed that, with exception of study No.2, all Credible Regions are wider than the corresponding Confidence Interval obtained on each study. The mean estimate of $\sigma_0$ is $0.418$ for both models, which is higher than most of the standard-error in $10$ studies (exceptions are studies \#2 and \#12), possibly generating the discrepancy between the length of intervals. 

(iii) We noticed study No. 2 has one of the smallest samples from all studies, here represented by $N$. \cite{GelmanEtAll} justifies that with  larger sample sizes (i.e., more than 50 persons in each treatment group in nearly all of the studies), the variances in each study are \textit{more precisely estimated} so the initial assumption that $S_2$ was the true standard error for study No. 2 may not be completely true and this might have impacted the determination of the estimates (this study has the highest difference between the posterior mean of $\delta_2$ and "delta[2]"). 

Another important aspect to notice is that Model-1 uses "Sediff[i]" to compose the variance of $Y$'s which might have interfered in our estimation for $\delta$ of each study. Using larger differences as the true $S_i$ imposes more variability of estimates on these studies, with consequent larger credible regions and larger differences between posterior mean of $\delta_i$ and "delta[i]".

\pagebreak

\begin{thebibliography}{7}

\bibitem{GelmanEtAll} 
Gelman, A., Carlin, J.B., Stern, H.S., Dunson, D.B., Vehtari, A., Rubin, D.B.
\textit{Bayesian Data Analysis}. pp.34 - CRC Press, 2014.

\bibitem{GelmanHill} 
Gelman, A., Hill, J.
\textit{Data Analysis using Regression and Multilevel/Hierarchical Models}, Cambridge Press, 2007.

\bibitem{Selvin} 
Selvin, S.
\textit{Practical Biostatistical Methods} - Dusbury Press, 1995

\bibitem{MarinRobert} 
Marin, J. M., Robert, C.
\textit{Bayesian Essentials with R}, 2nd. Ed. pp.49 - Springer, 2014.

\bibitem{Hirst}
Hirst J. A., Farmer A. J., Dyar A., Lung T. W. C., Stevens, R. J.
\textit{Estimating the effect of sulfonylurea on $HbA_{1c}$ in diabetes: a systematic review and meta-analysis}, Diabetologia 56:973984 (2013)

\bibitem{SuttonEtAll}
Sutton A.J., Abrams K.R., Jones D.R.,
\textit{Methods for Meta-Analysis in Medical Research}. London: Wiley, 2000.

\bibitem{FriendlyMeyer}
Friendly, M., Meyer., D.
\textit{Discrete Data Analysis with R - Visualization and Modeling Techniques for Categorical and Count Data}. CRC, 2013.
\end{thebibliography}

\pagebreak

\section{Appendix - R-Code}

```{r, echo=TRUE, eval=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, fig.width=5, fig.height=4.0)
library(tidyverse)
library(R2OpenBUGS)
library(kableExtra)
library(ggplot2)
library(forecast)

# The palette with black - Used in Graphs with :
cbp1 <- c("#999999", "#E69F00", "#56B4E9", "#009E73",
          "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
cbp2 <- c("#000000", "#E69F00", "#56B4E9", "#009E73",
          "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
cbp3 <- c("#FFDB6D", "#C4961A", "#F4EDCA", "#D16103", 
          "#C3D7A4", "#52854C", "#4E84C4", "#293352")

# Read database
bb_data <- read.csv("Data/brnbdy.csv")

# Introducing Scale Power transformation - Weisberg(2013) pp.189-190
scalePower <- function (Z,Lambda=0, Mod = FALSE) {
  # Introduced by Box, Cox (1964) for strictly positive Z
  if (Mod)
    gm = exp(sum(log(Z)/length(Z))) 
  else
    gm = 1.0
  if (Lambda == 0)
    return(gm*log(Z))
  else
    return(gm^(1-Lambda)*(Z^Lambda-1)/Lambda)
}

# Setup Transformed Variables
Y  <-  scalePower(bb_data$brain)
X  <-  scalePower(bb_data$body)


# Prepare dataframe and plot brain vs. body weight
df <- data.frame (lBrW = Y, lBdW = X)

# Plot Graph for 
df %>% 
  ggplot(aes(x=X, y=Y))+
  geom_point(size=1.8)+
  labs(x = "log(Body Weight)", y = "log(Brain Weight)") +
  ggtitle(NULL)+
  theme_bw()

# Remove Temporary Variable 
remove(df)

# Setup Initial Variables
Mu_alpha <- Mu_beta <- 0
Tau_alpha <- Tau_beta <- 0.0001
a_Tau<- b_Tau <- 0.0001  # v.16

# Setup Iteration Variables
N <- 20000

alpha <- beta <- tau <- rep(N,0)

# Setup \theta^{(0)}
alpha[1] <- beta[1] <- 0
tau[1] <- 1

# Calculates the value of parameters, given everything else
sample_Alpha <- function (X, Y, beta, tau, mu_alpha, tau_alpha) {
  tau_alphaStar <- length(Y)*tau+tau_alpha
  mu_alphaStar <- (tau*sum(Y-beta*X)+tau_alpha*mu_alpha)/tau_alphaStar
  alpha <- rnorm(1,mean=mu_alphaStar, sd=sqrt(1/tau_alphaStar))
  return(alpha)
}

sample_Beta <- function (X, Y, alpha, tau, mu_beta, tau_beta) {
  tau_betaStar<- tau*sum(X^2)+tau_beta
  mu_betaStar <- (tau*sum(X*(Y-alpha))+tau_beta*mu_beta)/tau_betaStar
  beta <- rnorm(1,mean=mu_betaStar, sd=sqrt(1/tau_betaStar))
  return(beta)
}

sample_Tau <- function (X, Y, alpha, beta, a_r, b_r) {
  a_rStar <- a_r+length(Y)/2
  b_rStar <- b_r+0.5*sum((Y-(alpha+beta*X))^2)
  tau <- rgamma(1,shape=a_rStar, rate=b_rStar)
  return(tau)
}

Gibbs <- function(X, Y, alpha, beta, tau, N) {
  for (i in 2:N) {
    alpha[i] <- sample_Alpha(X, Y, beta[i-1], tau[i-1], Mu_alpha, Tau_alpha)
    beta[i] <- sample_Beta(X, Y, alpha[i], tau[i-1], Mu_beta, Tau_beta)
    tau[i] <- sample_Tau(X, Y, alpha[i], beta[i], a_Tau, b_Tau)   # v.17
  }
  return(list(Alpha = alpha, Beta = beta, Tau = tau))
}

# Set Seed for reproducibility
set.seed(737)

#Calculates 3 chains
ThetaC1 <- Gibbs(X, Y, alpha, beta, tau, N)
ThetaC2 <- Gibbs(X, Y, alpha, beta, tau, N)
ThetaC3 <- Gibbs(X, Y, alpha, beta, tau, N)


# Plot ACF * ALPHA, BETA and TAU *

# Determine Burnin based on the convergence of ACF plot

# Build working dataframe
D_WorkAcfABT <- data.frame(ValCh1=ThetaC1$Alpha,
                           ValCh2=ThetaC1$Beta,
                           ValCh3=ThetaC1$Tau)


P1 <- ggAcf(D_WorkAcfABT$ValCh1, lag.max = 100)+
  labs(x = "Lag", y = expression(alpha)) +
  ggtitle(NULL)+
  theme_bw()
P2 <- ggAcf(D_WorkAcfABT$ValCh2, lag.max = 100)+
  labs(x = "Lag", y = expression(beta)) +
  ggtitle(NULL)+
  theme_bw()
P3 <- ggAcf(D_WorkAcfABT$ValCh3, lag.max = 100)+
  labs(x = "Lag", y = expression(tau)) +
  ggtitle(NULL)+
  theme_bw()

ggpubr::ggarrange(P1, P2, P3, ncol = 2, nrow = 2)

# Remove Working Variable to free memory
remove(D_WorkAcfABT)

# Set Burn-in parameter
Burnin <- 100

# Prepare Data-Frame for next steps
D1_Work <- rbind(data.frame(alpha=ThetaC1$Alpha[Burnin:N],
                            beta=ThetaC1$Beta[Burnin:N],
                            tau=ThetaC1$Tau[Burnin:N],Chain=factor(1)),
                 data.frame(alpha=ThetaC2$Alpha[Burnin:N],
                            beta=ThetaC2$Beta[Burnin:N],
                            tau=ThetaC2$Tau[Burnin:N],Chain=factor(2)),
                 data.frame(alpha=ThetaC3$Alpha[Burnin:N],
                            beta=ThetaC3$Beta[Burnin:N],
                            tau=ThetaC3$Tau[Burnin:N],Chain=factor(3)))


# Plot TracePlots * ALPHA, BETA and TAU *

set.seed(340)

Sz <- 5000

L <- N-Burnin+1

# Sampling 5000 points to generate "thinner" traceplots
S <- if (Sz <= L) sort(sample(Burnin:N, Sz, replace = FALSE)) else Burnin:N

# Build working dataframe
D_WorkTpAlpha <- data.frame(ValCh1=ThetaC1$Alpha[S],
                            ValCh2=ThetaC2$Alpha[S],
                            ValCh3=ThetaC3$Alpha[S])
P1 <- D_WorkTpAlpha %>%
  ggplot(aes(seq(from=1,to=Sz)))+
  geom_line(aes(y=ValCh1, colour=1), size=0.8)+ 
  geom_line(aes(y=ValCh2, colour=2), size=0.8)+ 
  geom_line(aes(y=ValCh3, colour=3), size=0.8)+ 
  labs(y = expression(alpha)) +
  theme_bw()

# Build working dataframe
D_WorkTpBeta <- data.frame(ValCh1=ThetaC1$Beta[S],
                           ValCh2=ThetaC2$Beta[S],
                           ValCh3=ThetaC3$Beta[S])
P2 <- D_WorkTpBeta %>%
  ggplot(aes(seq(from=1,to=Sz)))+
  geom_line(aes(y=ValCh1, colour=1), size=0.8)+ 
  geom_line(aes(y=ValCh2, colour=2), size=0.8)+ 
  geom_line(aes(y=ValCh3, colour=3), size=0.8)+ 
  labs(y = expression(beta)) +
  theme_bw()

# Build working dataframe
D_WorkTpTau <- data.frame(ValCh1=ThetaC1$Tau[S],
                          ValCh2=ThetaC2$Tau[S],
                          ValCh3=ThetaC3$Tau[S])
P3 <- D_WorkTpTau %>%
  ggplot(aes(seq(from=1,to=Sz)))+
  geom_line(aes(y=ValCh1, colour=1), size=0.8)+ 
  geom_line(aes(y=ValCh2, colour=2), size=0.8)+ 
  geom_line(aes(y=ValCh3, colour=3), size=0.8)+ 
  labs(y = expression(tau)) +
  theme_bw()

# Build working dataframe
D_WorkTpAlphaBeta <- data.frame(ValAlpha=c(ThetaC1$Alpha[S],
                                           ThetaC2$Alpha[S],
                                           ThetaC3$Alpha[S]),
                                ValBeta=c(ThetaC1$Beta[S],
                                          ThetaC2$Beta[S],
                                          ThetaC3$Beta[S]))

P4 <- D_WorkTpAlphaBeta %>%
  ggplot(aes(x=ValAlpha, y=ValBeta))+
  geom_point(size=0.6)+ 
  labs(x = expression(alpha), y = expression(beta)) +
  theme_bw()

ggpubr::ggarrange(P1+theme(axis.title.x=element_blank(),
                           legend.position="none"), 
                  P2+theme(axis.title.x=element_blank(),
                           legend.position="none"), 
                  P3+theme(axis.title.x=element_blank(),
                           legend.position="none"), 
                  P4+theme(legend.position="none"), ncol = 2, nrow = 2)

# Remove Working Variable to free memory
remove(D_WorkTpAlpha, D_WorkTpBeta, D_WorkTpTau, D_WorkTpAlphaBeta)

# Plot the Graphs of Alpha
P1 <- D1_Work %>%
  ggplot(mapping = aes(x = alpha, group = Chain))+
  geom_density(aes(colour=Chain), size=0.5)+
  labs(x = expression(alpha), y = "Density") +
  scale_color_manual(values = cbp2)+
  theme_bw()

# Plot the Graphs of Beta
P2 <- D1_Work %>%
  ggplot(mapping = aes(x = beta, group = Chain))+
  geom_density(aes(colour=Chain), size=0.5)+
  labs(x = expression(beta), y = "Density") +
  scale_color_manual(values = cbp2)+
  theme_bw()

# Plot the Graphs of Tau
P3 <-D1_Work %>%
  ggplot(mapping = aes(x = tau, group = Chain))+
  geom_density(aes(colour=Chain), size=0.5)+
  labs(x = expression(tau), y = "Density") +
  scale_color_manual(values = cbp2)+
  theme_bw()

ggpubr::ggarrange(P1, P2, P3,ncol = 1, nrow = 3)


# Summary Statistics - alpha
prt_SummaryStat <- function (sample_, parameter) {
  m_ <- mean(sample_)
  sd_ <- sd(sample_)
  cat("\n---- Summary Statistics for ",parameter," -----")
  cat("\nMean",parameter,": ",format(m_, digits = 2, nsmall = 3)) 
  cat("\nStd.Dev.",parameter,": ",format(sd_, digits = 2, nsmall = 4))
  CI <- qnorm(p = c(0.025,0.975), mean = m_, sd = sd_)
  # cat("\nC.I. for",parameter,": [",CI[1],",",CI[2],"]\n")
  cat("\n95% Credible Region for",parameter,"(", format(CI[1], digits = 2, nsmall = 3), 
      ",",format(CI[2], digits = 4, nsmall = 3),")\n")
}
prt_SummaryStat(D1_Work %>% pull(alpha), "alpha")
prt_SummaryStat(D1_Work %>% pull(beta), "beta")
prt_SummaryStat(D1_Work %>% pull(tau), "tau")


# Using the Alphas and Betas found on item (a) We will generate a sample of exp(Y_55)
D2_Work <- D1_Work %>% 
  mutate(Y_55 = alpha+beta*log(55))

D2_Work %>%
  ggplot(mapping = aes(x = exp(Y_55), group = Chain))+
  geom_density(aes(colour=Chain), size=0.8)+
  labs(x = expression("Estimated Brain Weight | Body Weight=55Kg"), y = "Density") +
  scale_color_manual(values = cbp2)+
  theme_bw()

# Summary Statistics for the Distribution of an animal with average bodyweight of 55
m_a <- mean(exp(D2_Work$Y_55))
sd_a <- sd(exp(D2_Work$Y_55))
cat("\n---- Summary Statistics for Brain Weight | Body Weight = 55kg -----")
cat("\nMean : ",format(m_a, digits = 2, nsmall = 3)) 
cat("\nStd.Dev. : ",format(sd_a, digits = 2, nsmall = 4))
CI_a <- qnorm(p = c(0.025,0.975), mean = m_a, sd = sd_a)
cat("\n95% Credible Region (", format(CI_a[1], digits = 2, nsmall = 3), 
    ",",format(CI_a[2], digits = 2, nsmall = 3),")\n")
cat("\n>>> From this model, our belief of Brain Weight for an animal with \n")
cat("    average Body Weight of 55Kg is",format(m_a, digits = 2, nsmall = 2),
    "grams, with a 95% credible\n")
cat("    region of (",format(CI_a[1], digits = 2, nsmall = 2),",",
    format(CI_a[2], digits = 2, nsmall = 2),") grams.\n")

# Setup Data-set - Read Data
shist_data <- read.table(file="data/SmokeAgeDeath.csv",header=TRUE,sep=",")
colnames(shist_data) <- c("Smoking", "Age", "Deaths", "PersonYears")

# Setup Variables
NSmok=nrow(shist_data)  # Number of Items in data-set
NCSmok=4   # Number of classes of Smoking
NCAge=5    # Number of Classes of Age

# Setup OpenBugs running parameters
NSim <- 30000   # No. of simulations for productio
NChain <- 5     # No. of chains for production
NThin <- 5      # n.thin parameter for production
Burnin <- 10000 # Burn-In parameter for production
Sz <- 5000      # Size of samples for trace/acf plots


# Printing the Data-Frame
shist_data %>%
  kbl(booktabs = TRUE,
      caption = "Deaths by Lung Cancer, per Age, Smoking Level and Person-Years") %>% 
  kable_styling(latex_options = "striped")


cat("
model{
for(i in 1:NSmok)
{
#Age   Smoking   Deaths  PersonYears

Deaths[i]~dpois(lambda[i])
log(lambda[i]) <-  log(PersonYears[i]) + beta0 + beta.Age[Age[i]]+ beta.Smoking[Smoking[i]] + bias[i] 
bias[i] ~dnorm(0,tau)
bias.adj[i] <- bias[i] - mean(bias[]) 
}

# Loop for Age
for(ia in 1:NCAge){         
beta.Age[ia]~dnorm(0,tau.Age)
beta.Age.adj[ia] <- beta.Age[ia] - mean(beta.Age[]) 
}

# Loop for Smoking Level
for(is in 1:NCSmok){         
beta.Smoking[is]~dnorm(0,tau.Smoking)
beta.Smoking.adj[is] <- beta.Smoking[is] - mean(beta.Smoking[])
}

# Initial Priors - Verify
beta0 ~ dnorm(0, 0.008)   # Overall rate of disease
beta0.adj <- beta0 + mean(bias[]) + mean(beta.Age[])+ mean(beta.Smoking[])

# Initial Priors - Verify
std ~ dunif(0, 4)
tau <- 1/pow(std,2)

# Initial Priors - Verify
std.Age ~dunif(0, 4)
tau.Age <- 1/pow(std.Age,2)
std.Smoking ~ dunif(0,4) 
tau.Smoking <- 1/pow(std.Smoking,2)

}", file="LungCancerCentered.txt")

# Setup Parameters
params=c("beta.Age.adj",  "std.Age", "beta.Smoking.adj", "std.Smoking", 
         "beta0.adj", "std", "tau", "tau.Age", "tau.Smoking")

# Setup Initial Values
init.fun=function(){list(
  beta.Age=rnorm(5), std.Age=runif(1,0,4),
  beta.Smoking=rnorm(4), std.Smoking=runif(1,0,4), 
  std=runif(1,0,4), beta0=rnorm(1),
  bias=rnorm(20,0,.1))}


# Run Open Bugs
set.seed(2602)
model.data <- list("Age", "Smoking","Deaths","PersonYears","NSmok", 
                   "NCSmok", "NCAge")
attach(shist_data)
LungCanc_v0=bugs(model.data, init.fun, params, model.file="LungCancerCentered.txt",
                 working.directory = ".", n.chains=NChain, n.iter=NSim, 
                 n.burnin=Burnin, n.thin=NThin)
detach(shist_data)


# Get Simulation from OpenBugs
SArray= LungCanc_v0$sims.array   # Data Arrays
vname=attr(SArray,"dimnames")[3][[1]]  # Variable Names


# Print Summary statistics of parameters oif Interest
RNames <- rownames(LungCanc_v0[["summary"]][,c(1:3,5,7)]) # List of parameters
df_Prt <- data.frame(Parameter = RNames)
df_Prt <- cbind(df_Prt, as_tibble(LungCanc_v0[["summary"]][,c(1:3,5,7)]))
df_Prt %>%
  kbl(booktabs = TRUE, 
      caption = "OpenBugs Summary - Deaths from Lung Cancer") %>% 
  kable_styling(latex_options = "striped")

# Plot Sample densities to Compare Distributions over *BETA-0*
D_WorkBeta0 <- as_tibble(data.frame(LvlBeta=as.vector(SArray[,,"beta0.adj"])))

D_WorkBeta0 %>%
  ggplot(aes(x = LvlStd))+
  geom_density(aes(x=LvlBeta), size=1.2)+ 
  labs(x = expression(list(beta[0])), y = "Density") +
  scale_color_manual(values = cbp2)+
  theme_bw()

# Remove Working Variable to free memory
remove(D_WorkBeta0)


# Plot Sample densities to Compare Distributions over *AGE*
D_WorkAge <- as_tibble(rbind(data.frame(LvlAge=as.vector(SArray[,,paste0("beta.Age.adj[", 1, "]")]),Age="0-45"),
                             data.frame(LvlAge=as.vector(SArray[,,paste0("beta.Age.adj[", 2, "]")]),Age="45-54"),
                             data.frame(LvlAge=as.vector(SArray[,,paste0("beta.Age.adj[", 3, "]")]),Age="55-64"),
                             data.frame(LvlAge=as.vector(SArray[,,paste0("beta.Age.adj[", 4, "]")]),Age="65-74"),
                             data.frame(LvlAge=as.vector(SArray[,,paste0("beta.Age.adj[", 5, "]")]),Age="74+")))


P0 <- D_WorkAge %>%
  ggplot(aes(x = LvlAge))+
  geom_density(aes(x=LvlAge,group=Age, colour=Age), size=1.2)+ 
  labs(x = expression(list(beta[Age_][i])), y = "Density") +
  scale_color_manual(values = cbp2)+
  theme_bw()

P0 + geom_segment(aes(x = -2, y = 2, xend = 1.5, yend = 4), size = 0.3,
                arrow = arrow(length = unit(0.5, "cm")))

# Remove Working Variable to free memory
remove(D_WorkAge)

# Plot Sample densities to Compare Distributions over *SMOKE LEVEL*
D_WorkSMoke <- as_tibble(rbind(data.frame(LvlSmoke=as.vector(SArray[,,paste0("beta.Smoking.adj[", 1, "]")]),Level="Lv1 - Never Smoked"),
                               data.frame(LvlSmoke=as.vector(SArray[,,paste0("beta.Smoking.adj[", 2, "]")]),Level="Lv2 - Past Smoker"),
                               data.frame(LvlSmoke=as.vector(SArray[,,paste0("beta.Smoking.adj[", 3, "]")]),Level="Lv3 - 1-20 Cig/Day"),
                               data.frame(LvlSmoke=as.vector(SArray[,,paste0("beta.Smoking.adj[", 4, "]")]),Level="Lv4 - +20 Cig/Day")))


D_WorkSMoke %>%
  ggplot(aes(x = LvlSmoke))+
  geom_density(aes(x=LvlSmoke,group=Level, colour=Level), size=1.2)+ 
  labs(x = expression(list(beta[Smoke_][i])), y = "Density") +
  scale_color_manual(values = cbp2)+
  theme_bw()

# Remove Working Variable to free memory
remove(D_WorkSMoke)

# Plot Sample densities to Compare Distributions over *STD DEVIATION* and *PRECISION*
D_WorkStd <- as_tibble(rbind(data.frame(LvlStd=as.vector(SArray[,,"std.Age"]),Level="Age"),
                             data.frame(LvlStd=as.vector(SArray[,,"std.Smoking"]),Level="Smoking")))

D_WorkPrc <- as_tibble(rbind(data.frame(LvlStd=as.vector(SArray[,,"tau.Age"]),Level="Age"),
                             data.frame(LvlStd=as.vector(SArray[,,"tau.Smoking"]),Level="Smoking")))


P1 <- D_WorkStd %>%
  ggplot(aes(x = LvlStd))+
  geom_density(aes(x=LvlStd,group=Level, colour=Level), size=1.2)+ 
  labs(x = expression(list(sigma[Age]~","~sigma[Smoke])), y = "Density") +
  scale_color_manual(values = cbp2)+
  theme_bw()

P2 <- D_WorkPrc %>%
  ggplot(aes(x = LvlStd))+
  geom_density(aes(x=LvlStd,group=Level, colour=Level), size=1.2)+ 
  labs(x = expression(list(tau[Age]~","~tau[Smoke])), y = "Density") +
  scale_color_manual(values = cbp2)+
  theme_bw()

ggpubr::ggarrange(P1, P2, ncol = 1, nrow = 2)

# Remove Working Variable to free memory
remove(D_WorkStd, D_WorkPrc)


# Plot TracePlots * BETA.AGE *

# Sampling 5000 points to generate "thinner" traceplots
set.seed(312)

L <- NSim-Burnin

S <- if (Sz <= L) sort(sample(1:L, Sz, replace = FALSE)) else 1:Sz

# Build working dataframe

P <- list()  

for (i in 1:5) {
  # Build working dataframe
  D_WorkTpAge <- data.frame(ValCh1=SArray[S,1,paste0("beta.Age.adj[", i, "]")],
                            ValCh2=SArray[S,2,paste0("beta.Age.adj[", i, "]")],
                            ValCh3=SArray[S,3,paste0("beta.Age.adj[", i, "]")],
                            ValCh4=SArray[S,4,paste0("beta.Age.adj[", i, "]")],
                            ValCh5=SArray[S,5,paste0("beta.Age.adj[", i, "]")])
  
  p <- D_WorkTpAge %>%
    ggplot(aes(seq(from=1,to=Sz)))+
    geom_line(aes(y=ValCh1, colour=1), size=0.8)+ 
    geom_line(aes(y=ValCh2, colour=2), size=0.8)+ 
    geom_line(aes(y=ValCh3, colour=3), size=0.8)+ 
    geom_line(aes(y=ValCh4, colour=4), size=0.8)+ 
    geom_line(aes(y=ValCh5, colour=5), size=0.8)+ 
    labs(y = eval(bquote(expression(beta[Age_][.(i)])))) +
    theme_bw()
  
  P <- c(P, list(p))
}

ggpubr::ggarrange(P[[1]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"), 
                  P[[2]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[3]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[4]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[5]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"), ncol = 3, nrow = 2)

# Remove Working Variable to free memory
remove(p, P, D_WorkTpAge)

# Plot TracePlots * BETA.SMOKING *

# Sampling 5000 points to generate "thinner" traceplots
set.seed(351)

L <- NSim-Burnin

S <- if (Sz <= L) sort(sample(1:L, Sz, replace = FALSE)) else 1:Sz

# Build working dataframe

P <- list()  

for (i in 1:4) {
  # Build working dataframe
  D_WorkTpSmok <- data.frame(ValCh1=SArray[S,1,paste0("beta.Smoking.adj[", i, "]")],
                             ValCh2=SArray[S,2,paste0("beta.Smoking.adj[", i, "]")],
                             ValCh3=SArray[S,3,paste0("beta.Smoking.adj[", i, "]")],
                             ValCh4=SArray[S,4,paste0("beta.Smoking.adj[", i, "]")],
                             ValCh5=SArray[S,5,paste0("beta.Smoking.adj[", i, "]")])
  
  p <- D_WorkTpSmok %>%
    ggplot(aes(seq(from=1,to=Sz)))+
    geom_line(aes(y=ValCh1, colour=1), size=0.8)+ 
    geom_line(aes(y=ValCh2, colour=2), size=0.8)+ 
    geom_line(aes(y=ValCh3, colour=3), size=0.8)+ 
    geom_line(aes(y=ValCh4, colour=4), size=0.8)+ 
    geom_line(aes(y=ValCh5, colour=5), size=0.8)+ 
    labs(y = eval(bquote(expression(beta[Smoke_][.(i)])))) +
    theme_bw()
  P <- c(P, list(p))
}

ggpubr::ggarrange(P[[1]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"), 
                  P[[2]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[3]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[4]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"), ncol = 2, nrow = 2)

# Remove Working Variable to free memory
remove(p, P, D_WorkTpSmok)

# Plot TracePlots * BETA.0 *

# Sampling 5000 points to generate "thinner" traceplots
set.seed(761)

L <- NSim-Burnin

S <- if (Sz <= L) sort(sample(1:L, Sz, replace = FALSE)) else 1:Sz

# Build working dataframe
D_WorkTpBeta0 <- data.frame(ValCh1=SArray[S,1,paste0("beta0.adj")],
                            ValCh2=SArray[S,2,paste0("beta0.adj")],
                            ValCh3=SArray[S,3,paste0("beta0.adj")],
                            ValCh4=SArray[S,4,paste0("beta0.adj")],
                            ValCh5=SArray[S,5,paste0("beta0.adj")])

P0 <- D_WorkTpBeta0 %>%
  ggplot(aes(seq(from=1,to=Sz)))+
  geom_line(aes(y=ValCh1, colour=1), size=0.8)+ 
  geom_line(aes(y=ValCh2, colour=2), size=0.8)+ 
  geom_line(aes(y=ValCh3, colour=3), size=0.8)+ 
  geom_line(aes(y=ValCh4, colour=4), size=0.8)+ 
  geom_line(aes(y=ValCh5, colour=5), size=0.8)+ 
  labs(x = "Iteration", y = expression(beta[0])) +
  theme_bw()

P0 + theme(axis.title.x=element_blank(),
           legend.position="none")

# Remove Working Variable to free memory
remove(D_WorkTpBeta0)

# Plot TracePlots * STD.AGE and STD.SMOKING *

# Sampling 5000 points to generate "thinner" traceplots
set.seed(390)

L <- NSim-Burnin

S <- if (Sz <= L) sort(sample(1:L, Sz, replace = FALSE)) else 1:Sz

# Build working dataframe
D_WorkTpStdAge <- data.frame(ValCh1=SArray[S,1,paste0("std.Age")],
                             ValCh2=SArray[S,2,paste0("std.Age")],
                             ValCh3=SArray[S,3,paste0("std.Age")],
                             ValCh4=SArray[S,4,paste0("std.Age")],
                             ValCh5=SArray[S,5,paste0("std.Age")])

P1 <- D_WorkTpStdAge %>%
  ggplot(aes(seq(from=1,to=Sz)))+
  geom_line(aes(y=ValCh1, colour=1), size=0.8)+ 
  geom_line(aes(y=ValCh2, colour=2), size=0.8)+ 
  geom_line(aes(y=ValCh3, colour=3), size=0.8)+ 
  geom_line(aes(y=ValCh4, colour=4), size=0.8)+ 
  geom_line(aes(y=ValCh5, colour=5), size=0.8)+ 
  labs(x = "Iteration", y = expression(sigma[Age])) +
  theme_bw()

# Build working dataframe
D_WorkTpStdSmok <- data.frame(ValCh1=SArray[S,1,paste0("std.Smoking")],
                              ValCh2=SArray[S,2,paste0("std.Smoking")],
                              ValCh3=SArray[S,3,paste0("std.Smoking")],
                              ValCh4=SArray[S,4,paste0("std.Smoking")],
                              ValCh5=SArray[S,5,paste0("std.Smoking")])

P2 <- D_WorkTpStdSmok %>%
  ggplot(aes(seq(from=1,to=Sz)))+
  geom_line(aes(y=ValCh1, colour=1), size=0.8)+ 
  geom_line(aes(y=ValCh2, colour=2), size=0.8)+ 
  geom_line(aes(y=ValCh3, colour=3), size=0.8)+ 
  geom_line(aes(y=ValCh4, colour=4), size=0.8)+ 
  geom_line(aes(y=ValCh5, colour=5), size=0.8)+ 
  labs(x = "Iteration", y = expression(sigma[Smoke])) +
  theme_bw()

ggpubr::ggarrange(P1+theme(axis.title.x=element_blank(),
                           legend.position="none"), 
                  P2+theme(axis.title.x=element_blank(),
                           legend.position="none"), ncol = 1, nrow = 2)

# Remove Working Variable to free memory
remove(D_WorkTpStdAge, D_WorkTpStdSmok)

# Plot TracePlots * BETA.AGE *

P <- list()  

for (i in 1:5) {
  # Build working dataframe
  D_WorkAcfAge <- data.frame(ValCh1=SArray[,1,paste0("beta.Age.adj[", i, "]")])
  
  p <- ggAcf(D_WorkAcfAge$ValCh1, lag.max = 100)+
    labs(x = "Lag", y = eval(bquote(expression(beta[Age_][.(i)])))) +
    ggtitle(NULL)+
    theme_bw()
  P <- c(P, list(p))
}

ggpubr::ggarrange(P[[1]]+theme(axis.title.x=element_blank(),
                               axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 8),
                               legend.position="none"), 
                  P[[2]]+theme(axis.title.x=element_blank(),
                               axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 8),
                               legend.position="none"),
                  P[[3]]+theme(axis.title.x=element_blank(),
                               axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 8),
                               legend.position="none"),
                  P[[4]]+theme(axis.title.x=element_blank(),
                               axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 8),
                               legend.position="none"),
                  P[[5]]+theme(axis.title.x=element_blank(),
                               axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 8),
                               legend.position="none"), ncol = 3, nrow = 2)

# Remove Working Variable to free memory
remove(p, P, D_WorkAcfAge)

# Plot ACF * BETA.SMOKING *

# Group 1-4 - Elaborate a List of Graphs

P <- list()  

for (i in 1:4) {
  # Build working dataframe
  D_WorkAcfSmok <- data.frame(ValCh1=SArray[,1,paste0("beta.Smoking.adj[", i, "]")])
  
  p <- ggAcf(D_WorkAcfSmok$ValCh1, lag.max = 100)+
    labs(x = "Lag", y = eval(bquote(expression(beta[Smoke_][.(i)])))) +
    ggtitle(NULL)+
    theme_bw()
  P <- c(P, list(p))
}

ggpubr::ggarrange(P[[1]]+theme(axis.title.x=element_blank(),
                               legend.position="none"), 
                  P[[2]]+theme(axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[3]]+theme(axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[4]]+theme(axis.title.x=element_blank(),
                               legend.position="none"), ncol = 2, nrow = 2)

# Remove Working Variable to free memory
remove(p, P, D_WorkAcfSmok)


# Plot TracePlots * BETA.0 *

# Build working dataframe
D_WorkAcfBeta0 <- data.frame(ValCh1=SArray[,1,paste0("beta0.adj")])

P1 <- ggAcf(D_WorkAcfBeta0$ValCh1, lag.max = 100)+
  labs(x = "Lag", y = expression(beta[0])) +
  ggtitle(NULL)+
  theme_bw()

P1+theme(axis.title.x=element_blank(),
         legend.position="none")

# Remove Working Variable to free memory
remove(D_WorkAcfBeta0)

# Comparison of Smoking Categories
SMokCat <- rbind(LungCanc_v0[["summary"]]["beta.Smoking.adj[1]",c(1:3,5,7)],
                 LungCanc_v0[["summary"]]["beta.Smoking.adj[4]",c(1:3,5,7)])
RNames <- c("beta.Smoking.adj[1]", "beta.Smoking.adj[4]")
df_Prt <- data.frame(Parameter = RNames)
df_Prt <- cbind(df_Prt, as_tibble(SMokCat))
df_Prt %>%
  kbl(booktabs = TRUE, 
      caption = "Smoking Categories - Never Smoked vs. >20 cigarettes per day") %>% 
  kable_styling(latex_options = "striped")

# Calculate Probability Increase between categories
ProbInc <- exp(SArray[,,"beta.Smoking.adj[4]"]-SArray[,,"beta.Smoking.adj[1]"])
m_Smoke <- mean(ProbInc)
sd_Smoke <- sd(ProbInc)
cat("\n---- Summary Statistics for Increase in Probability -----")
cat("\nMean : ",format(m_Smoke, digits = 2, nsmall = 3)) 
cat("\nStd.Dev. : ",format(sd_Smoke, digits = 2, nsmall = 4))
CI_Smoking <- qnorm(c(0.025,0.975), mean=m_Smoke, sd=sd_Smoke)
cat("\n95% Credible Region (", format(CI_Smoking[1], digits = 2, nsmall = 3), 
    ",",format(CI_Smoking[2], digits = 2, nsmall = 3),")\n")
cat("\n>>> From this model, our belief of increase probability of death by lung cancer\n")
cat("    of someone who smokes >20 cigarettes per day vs. who never smoked is",format(m_Smoke, digits = 2, nsmall = 2))
cat("\n    with corresponding 95% credible region of (",    format(CI_Smoking[1], digits = 2, nsmall = 2),",",
    format(CI_Smoking[2], digits = 2, nsmall = 2),")\n")

# Setup - Read Data Diabetics
diab_data <- read.table(file="data/DiabetesDrugEffect.csv",header=TRUE,sep=",")
NDiab <- nrow(diab_data)

# Setup OpenBugs running parameters

NSim <- 30000   # No. of simulations for productio
NChain <- 5     # No. of chains for production
NThin <- 5      # n.thin parameter for production
Burnin <- 10000 # Burn-In parameter for production
Sz <- 5000      # Size of samples for trace/acf plots


### **************** M O D E L  -  1 ****************** ###

### ---- SETUP MODELS ---- ###

# MODEL 1 - Splitted
cat("
model{
for(i in 1:NDiab)
{
#StudyID  StudyN  diff  Sediff

diff[i]~dnorm(delta[i], tau[i]) #v.19
delta[i]~dnorm(theta, tau0)
tau[i] <- 1/pow(Sediff[i],2)

}

# Initial Priors - Verify
theta ~dnorm(-1.5, 30)  # v.19
sigma0 ~dgamma(2, 1)    # v.19
tau0 <- 1/pow(sigma0,2)

}", file="DiabetDrug_M1.txt")

### ---- PARAM SETTINGS ---- ###

# Setup Parameters
paramsM1=c("theta", "tau0", "sigma0", "delta")

### ---- INITIALIZATION PROCEDURES ---- ###

# Setup Initial Values for Model 1
init.funM1=function(){list(
  theta = rnorm(1, mean=-1.5, sd=sqrt(.3)),
  sigma0=rgamma(1, 2, 1))}  # v.19

### ---- PROCESS MODELS ON OPENBUGS ---- ###

# Run Open Bugs - Model 1
set.seed(2602)
model1.data <- list("StudyID", "StudyN", "diff", "Sediff","NDiab")
attach(diab_data)
DiabetDrug_M1=bugs(model1.data, init.funM1, paramsM1, model.file="DiabetDrug_M1.txt",
                   working.directory = ".", n.chains=NChain, n.iter=NSim, n.burnin=Burnin, n.thin=NThin)
detach(diab_data)

### ---- ORGANIZE OUTPUTS ---- ###

# Get Simulation from OpenBugs - Model 1
DArrayM1= DiabetDrug_M1$sims.array   # Data Arrays
vname=attr(DArrayM1,"dimnames")[3][[1]]  # Variable Names


### **************** M O D E L  -  2 ****************** ###

### ---- SETUP MODELS ---- ###

cat("
model{
for(i in 1:NDiab)
{
#StudyID  StudyN  diff  Sediff

diff[i]~dnorm(theta, tau[i])
tau[i] <- 1/(pow(sigma0,2)+pow(Sediff[i],2))
}

# Initial Priors - Verify
theta ~dnorm(-1.5, 30)  # v.19
sigma0 ~dgamma(2, 1)    # v.19

}", file="DiabetDrug_M2.txt")

### ---- PARAM SETTINGS ---- ###

# Setup Parameters
paramsM2=c("theta", "sigma0")


### ---- INITIALIZATION PROCEDURES ---- ###

# Setup Initial Values for Model 2
init.funM2=function(){list(
  theta = rnorm(1, mean=-1.5, sd=sqrt(.3)),
  sigma0=rgamma(1, 2, 1))}  # v.19

### ---- PROCESS MODELS ON OPENBUGS ---- ###

# Run Open Bugs - Model 2
set.seed(8421)
model2.data <- list("StudyID", "StudyN", "diff", "Sediff","NDiab")
attach(diab_data)
DiabetDrug_M2=bugs(model2.data, init.funM2, paramsM2, model.file="DiabetDrug_M2.txt",
                   working.directory = ".", n.chains=NChain, n.iter=NSim, n.burnin=Burnin, n.thin=NThin)
detach(diab_data)

### ---- ORGANIZE OUTPUTS ---- ###

# Get Simulation from OpenBugs - Model 2
DArrayM2= DiabetDrug_M2$sims.array   # Data Arrays
vname=attr(DArrayM2,"dimnames")[3][[1]]  # Variable Names


# Print Summary statistics of parameters oif Interest
RNames <- rownames(DiabetDrug_M1[["summary"]][,c(1:3,5,7)]) # List of parameters
df_Prt <- data.frame(Parameter = RNames)
df_Prt <- cbind(df_Prt, as_tibble(DiabetDrug_M1[["summary"]][,c(1:3,5,7)]))
df_Prt %>%
  kbl(booktabs = TRUE, 
      caption = "Model 1 - Difference between Drug A and B") %>% 
  kable_styling(latex_options = "striped")


# Print Summary statistics of parameters oif Interest
RNames <- rownames(DiabetDrug_M2[["summary"]][,c(1:3,5,7)]) # List of parameters
df_Prt <- data.frame(Parameter = RNames)
df_Prt <- cbind(df_Prt, as_tibble(DiabetDrug_M2[["summary"]][,c(1:3,5,7)]))
df_Prt %>%
  kbl(booktabs = TRUE, 
      caption = "Model 2 - Difference between Drug A and B") %>% 
  kable_styling(latex_options = "striped")

# Plot TracePlots and ACF of *THETA *

# Sampling 5000 points to generate "thinner" traceplots
set.seed(3961)

L <- NSim - Burnin

S <- if (Sz <= L) sort(sample(1:L, Sz, replace = FALSE)) else 1:Sz

# Build working dataframe
D_WorkTpTheta <- data.frame(ValCh1=DArrayM1[S,1,paste0("theta")],
                            ValCh2=DArrayM1[S,2,paste0("theta")],
                            ValCh3=DArrayM1[S,3,paste0("theta")],
                            ValCh4=DArrayM1[S,4,paste0("theta")],
                            ValCh5=DArrayM1[S,5,paste0("theta")])

P1 <- D_WorkTpTheta %>%
  ggplot(aes(seq(from=1,to=Sz)))+
  geom_line(aes(y=ValCh1, colour=1), size=0.8)+ 
  geom_line(aes(y=ValCh2, colour=2), size=0.8)+ 
  geom_line(aes(y=ValCh3, colour=3), size=0.8)+ 
  geom_line(aes(y=ValCh4, colour=4), size=0.8)+ 
  geom_line(aes(y=ValCh5, colour=5), size=0.8)+ 
  labs(y = expression(theta)) +
  theme_bw()

P2 <- ggAcf(DArrayM1[,1,paste0("theta")], lag.max = 100)+
  labs(x = "Lag", y = expression(theta)) +
  ggtitle(NULL)+
  theme_bw()

ggpubr::ggarrange(P1+theme(axis.title.x=element_blank(),
                           legend.position="none"), P2, ncol = 1, nrow = 2)

# Remove Working Variable to free memory
remove(D_WorkTpTheta)

# Plot TracePlots * SIGMA0*

# Sampling 5000 points to generate "thinner" traceplots
set.seed(2391)

L <- NSim - Burnin

S <- if (Sz <= L) sort(sample(1:L, Sz, replace = FALSE)) else 1:Sz

# Build working dataframe
D_WorkTpSigma0 <- data.frame(ValCh1=DArrayM1[S,1,paste0("sigma0")],
                             ValCh2=DArrayM1[S,2,paste0("sigma0")],
                             ValCh3=DArrayM1[S,3,paste0("sigma0")],
                             ValCh4=DArrayM1[S,4,paste0("sigma0")],
                             ValCh5=DArrayM1[S,5,paste0("sigma0")])

P1 <- D_WorkTpSigma0 %>%
  ggplot(aes(seq(from=1,to=Sz)))+
  geom_line(aes(y=ValCh1, colour=1), size=0.8)+ 
  geom_line(aes(y=ValCh2, colour=2), size=0.8)+ 
  geom_line(aes(y=ValCh3, colour=3), size=0.8)+ 
  geom_line(aes(y=ValCh4, colour=4), size=0.8)+ 
  geom_line(aes(y=ValCh5, colour=5), size=0.8)+ 
  labs(y = expression(sigma[0])) +
  theme_bw()

P2 <- ggAcf(DArrayM1[,1,paste0("sigma0")], lag.max = 100)+
  labs(x = "Lag", y = expression(sigma[0])) +
  ggtitle(NULL)+
  theme_bw()

ggpubr::ggarrange(P1+theme(axis.title.x=element_blank(),
                           legend.position="none"), P2, ncol = 1, nrow = 2)

# Remove Working Variable to free memory
remove(D_WorkTpSigma0)

# Plot TracePlots for Deltas

# Sampling 5000 points to generate "thinner" traceplots
set.seed(746)

L <- NSim - Burnin

S <- if (Sz <= L) sort(sample(1:L, Sz, replace = FALSE)) else 1:Sz

# Group 1-4 - Elaborate a List of Graphs

P <- list()  

for (i in 1:4) {
  # Build working dataframe
  D_WorkTpDelta <- data.frame(ValCh1=DArrayM1[S,1,paste0("delta[", i, "]")],
                              ValCh2=DArrayM1[S,2,paste0("delta[", i, "]")],
                              ValCh3=DArrayM1[S,3,paste0("delta[", i, "]")],
                              ValCh4=DArrayM1[S,4,paste0("delta[", i, "]")],
                              ValCh5=DArrayM1[S,5,paste0("delta[", i, "]")])
  p <- D_WorkTpDelta %>%
    ggplot(aes(seq(from=1,to=Sz)))+
    geom_line(aes(y=ValCh1, colour=1), size=0.8)+ 
    geom_line(aes(y=ValCh2, colour=2), size=0.8)+ 
    geom_line(aes(y=ValCh3, colour=3), size=0.8)+ 
    geom_line(aes(y=ValCh4, colour=4), size=0.8)+ 
    geom_line(aes(y=ValCh5, colour=5), size=0.8)+ 
    labs(y = eval(bquote(expression(delta[.(i)])))) +
    theme_bw()
  P <- c(P, list(p))
}

ggpubr::ggarrange(P[[1]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"), 
                  P[[2]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[3]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[4]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"), ncol = 2, nrow = 2)

# Remove Working Variable to free memory
remove(p, P, D_WorkTpDelta)

# Plot TracePlots for Deltas

# Sampling 5000 points to generate "thinner" traceplots
set.seed(456)

L <- NSim - Burnin

S <- if (Sz <= L) sort(sample(1:L, Sz, replace = FALSE)) else 1:Sz

# Group 5-8 - Elaborate a List of Graphs

P <- list()  

for (i in 5:8) {
  # Build working dataframe
  D_WorkTpDelta <- data.frame(ValCh1=DArrayM1[S,1,paste0("delta[", i, "]")],
                              ValCh2=DArrayM1[S,2,paste0("delta[", i, "]")],
                              ValCh3=DArrayM1[S,3,paste0("delta[", i, "]")],
                              ValCh4=DArrayM1[S,4,paste0("delta[", i, "]")],
                              ValCh5=DArrayM1[S,5,paste0("delta[", i, "]")])
  p <- D_WorkTpDelta %>%
    ggplot(aes(seq(from=1,to=Sz)))+
    geom_line(aes(y=ValCh1, colour=1), size=0.8)+ 
    geom_line(aes(y=ValCh2, colour=2), size=0.8)+ 
    geom_line(aes(y=ValCh3, colour=3), size=0.8)+ 
    geom_line(aes(y=ValCh4, colour=4), size=0.8)+ 
    geom_line(aes(y=ValCh5, colour=5), size=0.8)+ 
    labs(y = eval(bquote(expression(delta[.(i)])))) +
    theme_bw()
  P <- c(P, list(p))
}

ggpubr::ggarrange(P[[1]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"), 
                  P[[2]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[3]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[4]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"), ncol = 2, nrow = 2)

# Remove Working Variable to free memory
remove(p, P, D_WorkTpDelta)

# Plot TracePlots for Deltas

# Sampling 5000 points to generate "thinner" traceplots
set.seed(198)

L <- NSim - Burnin

S <- if (Sz <= L) sort(sample(1:L, Sz, replace = FALSE)) else 1:Sz

# Group 9-12 - Elaborate a List of Graphs

P <- list()  

for (i in 9:12) {
  # Build working dataframe
  D_WorkTpDelta <- data.frame(ValCh1=DArrayM1[S,1,paste0("delta[", i, "]")],
                              ValCh2=DArrayM1[S,2,paste0("delta[", i, "]")],
                              ValCh3=DArrayM1[S,3,paste0("delta[", i, "]")],
                              ValCh4=DArrayM1[S,4,paste0("delta[", i, "]")],
                              ValCh5=DArrayM1[S,5,paste0("delta[", i, "]")])
  p <- D_WorkTpDelta %>%
    ggplot(aes(seq(from=1,to=Sz)))+
    geom_line(aes(y=ValCh1, colour=1), size=0.8)+ 
    geom_line(aes(y=ValCh2, colour=2), size=0.8)+ 
    geom_line(aes(y=ValCh3, colour=3), size=0.8)+ 
    geom_line(aes(y=ValCh4, colour=4), size=0.8)+ 
    geom_line(aes(y=ValCh5, colour=5), size=0.8)+ 
    labs(y = eval(bquote(expression(delta[.(i)])))) +
    theme_bw()
  P <- c(P, list(p))
}

ggpubr::ggarrange(P[[1]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"), 
                  P[[2]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[3]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[4]]+theme(axis.text.x=element_blank(),
                               axis.title.x=element_blank(),
                               legend.position="none"), ncol = 2, nrow = 2)

# Remove Working Variable to free memory
remove(p, P, D_WorkTpDelta)

# Plot ACF plots for Deltas

# Sampling 5000 points to generate "thinner" traceplots

# Group 1-4 - Elaborate a List of Graphs

P <- list()  

for (i in 1:4) {
  # Build working dataframe
  D_WorkTpDelta <- data.frame(ValCh1=DArrayM1[,1,paste0("delta[", i, "]")])
  
  p <- ggAcf(D_WorkTpDelta$ValCh1, lag.max = 100)+
    labs(x = "Lag", y = eval(bquote(expression(delta[.(i)])))) +
    ggtitle(NULL)+
    theme_bw()
  P <- c(P, list(p))
}

ggpubr::ggarrange(P[[1]]+theme(axis.title.x=element_blank(),
                               legend.position="none"), 
                  P[[2]]+theme(axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[3]]+theme(axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[4]]+theme(axis.title.x=element_blank(),
                               legend.position="none"), ncol = 2, nrow = 2)

# Remove Working Variable to free memory
remove(p, P, D_WorkTpDelta)

# Plot ACF plots for Deltas

# Sampling 5000 points to generate "thinner" traceplots

# Group 5-8 - Elaborate a List of Graphs

P <- list()  

for (i in 5:8) {
  # Build working dataframe
  D_WorkTpDelta <- data.frame(ValCh1=DArrayM1[,1,paste0("delta[", i, "]")])
  
  p <- ggAcf(D_WorkTpDelta$ValCh1, lag.max = 100)+
    labs(x = "Lag", y = eval(bquote(expression(delta[.(i)])))) +
    ggtitle(NULL)+
    theme_bw()
  P <- c(P, list(p))
}

ggpubr::ggarrange(P[[1]]+theme(axis.title.x=element_blank(),
                               legend.position="none"), 
                  P[[2]]+theme(axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[3]]+theme(axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[4]]+theme(axis.title.x=element_blank(),
                               legend.position="none"), ncol = 2, nrow = 2)

# Remove Working Variable to free memory
remove(p, P, D_WorkTpDelta)

# Plot ACF plots for Deltas

# Group 9-12 - Elaborate a List of Graphs

P <- list()  

for (i in 9:12) {
  # Build working dataframe
  D_WorkTpDelta <- data.frame(ValCh1=DArrayM1[,1,paste0("delta[", i, "]")])
  
  p <- ggAcf(D_WorkTpDelta$ValCh1, lag.max = 100)+
    labs(x = "Lag", y = eval(bquote(expression(delta[.(i)])))) +
    ggtitle(NULL)+
    theme_bw()
  P <- c(P, list(p))
}

ggpubr::ggarrange(P[[1]]+theme(axis.title.x=element_blank(),
                               legend.position="none"), 
                  P[[2]]+theme(axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[3]]+theme(axis.title.x=element_blank(),
                               legend.position="none"),
                  P[[4]]+theme(axis.title.x=element_blank(),
                               legend.position="none"), ncol = 2, nrow = 2)
# Remove Working Variable to free memory
remove(p, P, D_WorkTpDelta)

# Plot TracePlots and ACF of *THETA *

# Sampling 5000 points to generate "thinner" traceplots
set.seed(3871)

L <- NSim - Burnin

S <- if (Sz <= L) sort(sample(1:L, Sz, replace = FALSE)) else 1:Sz

# Build working dataframe
D_WorkTpTheta <- data.frame(ValCh1=DArrayM2[S,1,paste0("theta")],
                            ValCh2=DArrayM2[S,2,paste0("theta")],
                            ValCh3=DArrayM2[S,3,paste0("theta")],
                            ValCh4=DArrayM2[S,4,paste0("theta")],
                            ValCh5=DArrayM2[S,5,paste0("theta")])

P1 <- D_WorkTpTheta %>%
  ggplot(aes(seq(from=1,to=Sz)))+
  geom_line(aes(y=ValCh1, colour=1), size=0.8)+ 
  geom_line(aes(y=ValCh2, colour=2), size=0.8)+ 
  geom_line(aes(y=ValCh3, colour=3), size=0.8)+ 
  geom_line(aes(y=ValCh4, colour=4), size=0.8)+ 
  geom_line(aes(y=ValCh5, colour=5), size=0.8)+ 
  labs(y = expression(theta)) +
  theme_bw()

P2 <- ggAcf(DArrayM2[,1,paste0("theta")], lag.max = 100)+
  labs(x = "Lag", y = expression(theta)) +
  ggtitle(NULL)+
  theme_bw()

ggpubr::ggarrange(P1+theme(axis.title.x=element_blank(),
                           legend.position="none"), P2, ncol = 1, nrow = 2)

# Remove Working Variable to free memory
remove(D_WorkTpTheta)

# Plot TracePlots * SIGMA0*

# Sampling 5000 points to generate "thinner" traceplots
set.seed(8463)

L <- NSim - Burnin

S <- if (Sz <= L) sort(sample(1:L, Sz, replace = FALSE)) else 1:Sz

# Build working dataframe
D_WorkTpSigma0 <- data.frame(ValCh1=DArrayM2[S,1,paste0("sigma0")],
                             ValCh2=DArrayM2[S,2,paste0("sigma0")],
                             ValCh3=DArrayM2[S,3,paste0("sigma0")],
                             ValCh4=DArrayM2[S,4,paste0("sigma0")],
                             ValCh5=DArrayM2[S,5,paste0("sigma0")])

P1 <- D_WorkTpSigma0 %>%
  ggplot(aes(seq(from=1,to=Sz)))+
  geom_line(aes(y=ValCh1, colour=1), size=0.8)+ 
  geom_line(aes(y=ValCh2, colour=2), size=0.8)+ 
  geom_line(aes(y=ValCh3, colour=3), size=0.8)+ 
  geom_line(aes(y=ValCh4, colour=4), size=0.8)+ 
  geom_line(aes(y=ValCh5, colour=5), size=0.8)+ 
  labs(y = expression(sigma[0])) +
  theme_bw()

P2 <- ggAcf(DArrayM2[,1,paste0("sigma0")], lag.max = 100)+
  labs(x = "Lag", y = expression(sigma[0])) +
  ggtitle(NULL)+
  theme_bw()

ggpubr::ggarrange(P1+theme(axis.title.x=element_blank(),
                           legend.position="none"), P2, ncol = 1, nrow = 2)

# Remove Working Variable to free memory
remove(D_WorkTpSigma0)

# Plot Sample densities to Compare Distributions of *THETA* and *SIGMA0*

set.seed(9023)

L <- NSim-Burnin

S <- if (Sz <= L) sort(sample(1:L, Sz, replace = FALSE)) else 1:Sz

D_WorkThetas <- as_tibble(rbind(data.frame(LvlStd=as.vector(DArrayM1[S,,paste0("theta")]), Model = "M1"),
                                data.frame(LvlStd=as.vector(DArrayM2[S,,paste0("theta")]), Model = "M2")))
D_WorkSigmas <- as_tibble(rbind(data.frame(LvlStd=as.vector(DArrayM1[S,,paste0("sigma0")]), Model = "M1"),
                                data.frame(LvlStd=as.vector(DArrayM2[S,,paste0("sigma0")]), Model = "M2")))


P1 <- D_WorkThetas %>%
  ggplot(aes(x = LvlStd))+
  geom_density(aes(x=LvlStd, group=Model, colour=Model), size=0.8)+ 
  labs(x = expression(list(theta[Model_][1]~"vs."~theta[Model_][2])), y = "Density") +
  scale_color_manual(values = cbp2)+
  theme_bw()

P2 <- D_WorkSigmas %>%
  ggplot(aes(x = LvlStd))+
  geom_density(aes(x=LvlStd, group=Model, colour=Model), size=0.8)+ 
  labs(x = expression(list((sigma[0])[Model_][1]~"vs."~(sigma[0])[Model_][2])), y = "Density") +
  scale_color_manual(values = cbp2)+
  theme_bw()

ggpubr::ggarrange(P1, P2, ncol = 1, nrow = 2)

# Remove Working Variable to free memory
remove(D_WorkThetas, D_WorkSigmas)


# Prepare Data - Get delta[1-12]
D3_Work <- as.data.frame(DiabetDrug_M1[["summary"]][4:15,c(1:3,5,7)])

# Generates Studies labels
StudyLst <- rapply(list(diab_data$StudyID), sprintf, fmt = "%02d", how = "replace")

# Collects C.I's of observed values from studies
CI_Study <- data.frame(ID = StudyLst[[1]],   
                       lower = diab_data$diff-diab_data$Sediff, 
                       estim = diab_data$diff, 
                       upper = diab_data$diff+diab_data$Sediff, 
                       Source = "Study")

# Collects C.R's from Summary report 
CI_Estim <- data.frame(ID = StudyLst[[1]], 
                       lower = D3_Work$`2.5%`, 
                       estim = D3_Work$mean, 
                       upper = D3_Work$`97.5%`, 
                       Source = "Model-1")

# Organize data to generate graphs & analysis
D4_Work <- rbind(CI_Study, CI_Estim)

#  (i) Identifying the direction of change of estimates
# (ii) Identifying differences o length of CR and CI for delta and "diff"
P1 <- D4_Work %>% 
  ggplot() +
  geom_point(aes(x=ID, y=estim, color = Source, shape = Source, size = 0.5)) +
  geom_errorbar(aes(x=ID, ymin=lower, ymax=upper,
                    color = Source), width = 1) +
  xlab("Study No.") +
  ylab(expression("C.R.+"~delta[i]~" and 'diff'")) +
  scale_color_manual(values = cbp2)+
  facet_wrap(~Source)+
  theme_bw()

P1 + theme(legend.position="none",
           axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 8))

# (i) Identifying the direction of posterior mean of delta and "diff"
P2 <- D4_Work %>% 
  ggplot() +
  geom_point(aes(x=ID, y=estim, colour = Source, shape = Source), size=5) +
  xlab(expression("Study No.")) +
  ylab(expression(delta[i]~" and 'diff'")) +
  scale_color_manual(values = cbp2)+
  theme_bw()

P2 + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 8))

remove(D3_Work, D4_Work, StudyLst)

```






